\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Algebra}{2}

\begin{document}

\SpeedTitleBlock{Daily Algebra Drill \#2}{David Akinyele-Aje}
\noindent\textit{New toolkit today: Vieta's formulas for cubics $\cdot$ Nested surds $\cdot$ Telescoping products $\cdot$ Lagrange/Brahmagupta identity $\cdot$ Factor theorem $\cdot$ Cyclic sums under constraints.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15\,s each)}}
%═══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Factorise, simplify, or evaluate instantly. No expansion.}

\begin{enumerate}

\item Factorise completely: $\;x^3 - 27$.

\item Without expanding, evaluate $\;({\sqrt{5}+\sqrt{3}})({\sqrt{5}-\sqrt{3}})$.

\item Without a calculator, evaluate $\;999^2 - 1$.

\item Given $a - b = 4$ and $ab = 5$, find $a^2 + b^2$.

\item Expand and simplify: $\;(x+y)^3 - x^3 - y^3$.

\item Factorise completely: $\;3x^2 - 75$.

\item Rationalise the denominator: $\;\dfrac{1}{\sqrt{3}+\sqrt{2}}$.

\item Factorise: $\;x^2(x-1)-(x-1)$.

\item The roots of $x^2 - 5x + 3 = 0$ are $p$ and $q$. Find $p^2 + q^2$.

\item Simplify: $\;\dfrac{2^{n+3}-2^n}{2^n}$.

\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50\,s each)}}
%═══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Fractions, rearranging, hidden factorisations. Elegance over brute force.}

\begin{enumerate}

\item Simplify: $\;\dfrac{x}{x^2-1}+\dfrac{1}{x-1}$.

\item Factorise: $\;x^4 - x^2 - 12$.

\item Make $r$ the subject of $\;S = \dfrac{a}{1-r}$.

\item Simplify: $\;\left(1+\dfrac{1}{x}\right)\!\left(1-\dfrac{1}{x+1}\right)$.

\item Given $x+y=3$ and $xy=1$, find $x^3+y^3$.

\item Simplify: $\;\dfrac{(a+b)^2-(a-b)^2}{4ab}$.

\item Factorise: $\;x^3+3x^2-x-3$.

\item Solve: $\;\dfrac{x-1}{x+2}=\dfrac{x+3}{x+8}$.

\item Simplify: $\;\dfrac{\sqrt{a}-\sqrt{b}}{\sqrt{a}+\sqrt{b}}+\dfrac{\sqrt{a}+\sqrt{b}}{\sqrt{a}-\sqrt{b}}$.

\item The cubic $x^3 - 6x^2 + 11x - 6 = 0$ has roots $\alpha, \beta, \gamma$.
      Without solving it, find $\alpha^2+\beta^2+\gamma^2$ and $\alpha\beta\gamma$.

\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90\,s each)}}
%═══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Find structure before computing. Substitution is the key, not expansion.}

\begin{enumerate}

\item Solve: $\;x^4-13x^2+36=0$.

\item Solve: $\;(x+1)(x+2)(x+3)(x+4)=24$.

\item Simplify: $\;\sqrt{6+2\sqrt{5}}$.

\item Given $a+b+c=1$ and $a^2+b^2+c^2=1$, find the value of $ab+bc+ca$
      and hence evaluate $a^3+b^3+c^3-3abc$.

\item Solve: $\;(x-1)(x-3)(x-5)(x-7)+15=0$.

\item If $x=2+\sqrt{3}$, find $x^2+\dfrac{1}{x^2}$ without expanding $x^2$ directly.

\item Given $\dfrac{a}{b}+\dfrac{b}{a}=3$, find $\dfrac{a^2}{b^2}+\dfrac{b^2}{a^2}$.

\item If $x+y+z=0$, simplify $\;\dfrac{(x^2+y^2+z^2)^2}{x^4+y^4+z^4}$.

\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(TMUA / SMC difficulty)}}
%═══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Each has a solution under five lines. Find the insight, not the computation.}

\begin{enumerate}

\item Evaluate:
\[
  \frac{1}{1\cdot2\cdot3}+\frac{1}{2\cdot3\cdot4}+\frac{1}{3\cdot4\cdot5}
    +\cdots+\frac{1}{8\cdot9\cdot10}.
\]

\item The roots of $x^2+px+q=0$ are $r$ and $s$.
      Find, in terms of $p$ and $q$, the monic quadratic whose roots are
      $r^2+s$ and $s^2+r$.

\item The equation $\;\dfrac{x}{y+z}=\dfrac{y}{x+z}=\dfrac{z}{x+y}=k$\;
      holds for real $x,y,z$ with $x+y+z\neq0$.
      Find all possible values of $k$.

\item Prove the \textbf{Brahmagupta--Fibonacci identity}:
\[
  (a^2+b^2)(c^2+d^2)=(ac+bd)^2+(ad-bc)^2.
\]
Hence express $85\times 65$ as a sum of two perfect squares in two different ways.

\item Find all integers $n\geq 1$ for which $\;n^2+n+1\;$ divides $\;n^3-1$.

\end{enumerate}

\SpeedClosing{``An elegant solution is not luck --- it is the reward for seeing the structure
    that was always there.''}

\end{document}
