\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Combinatorics}{6}

\begin{document}

\SpeedTitleBlock{Daily Combinatorics Drill \#6 --- Answers \& Investigations}
\noindent\textit{New toolkit today: Pigeonhole Principle (basic \& strong) $\cdot$ Worst-Case Thinking $\cdot$ Double Counting $\cdot$ The Handshake Lemma.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15\,s each)}}
%═══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}

%── A1 ──────────────────────────────────────────────────────────────────────────
\item Socks to guarantee a matching pair?

\ans{$3$}
\method{Worst case: one of each colour (2 socks). One more forces a match.}
\inv{The answer is (number of colours) $+1$, independent of how many socks the drawer holds. What DOES the drawer size affect? (Whether the guarantee is achievable at all --- with one black sock total, no black pair exists.)}

%── A2 ──────────────────────────────────────────────────────────────────────────
\item Draws to guarantee a black pair?

\ans{$12$}
\method{Worst case: all 10 white socks first, then 2 black.}
\inv{``A pair'' vs ``a \emph{black} pair'' changed the answer from 3 to 12 --- adversarial drawing exhausts everything unhelpful first. Rule: to guarantee $X$, assume the drawer fights you optimally, then add what $X$ still needs.}

%── A3 ──────────────────────────────────────────────────────────────────────────
\item 13 people, two share a birth month?

\ans{True}
\method{13 pigeons, 12 month-boxes.}
\inv{The principle is an \emph{existence} claim: it never says which month. That non-constructive character is what makes pigeonhole proofs feel like magic --- and what D-section write-ups must state cleanly.}

%── A4 ──────────────────────────────────────────────────────────────────────────
\item Guarantee two share a weekday of birth?

\ans{$8$}
\method{7 boxes, so 8 people.}
\inv{General basic form: $k$ boxes need $k+1$ pigeons. Say it as a sentence: ``if $k+1$ objects go into $k$ boxes, some box holds two.'' Being able to \emph{state} the principle precisely is worth marks in proofs.}

%── A5 ──────────────────────────────────────────────────────────────────────────
\item Integers to guarantee two of the same parity?

\ans{$3$}
\method{Two parity boxes.}
\inv{Same-parity difference is even --- so among any 3 integers, two differ by an even number (A10 is the 4-integer version --- why does it also hold with just 3?).}

%── A6 ──────────────────────────────────────────────────────────────────────────
\item Handshake counts total 18: how many handshakes?

\ans{$9$}
\method{Each handshake is counted twice, once by each participant: $18/2$.}
\inv{The Handshake Lemma: $\sum\deg=2\cdot(\text{edges})$. Immediate corollary: the number of people with an ODD handshake count is even. Check it at your next gathering; prove it from the lemma in one line.}

%── A7 ──────────────────────────────────────────────────────────────────────────
\item Row totals sum $=$ column totals sum?

\ans{True}
\method{Both equal the sum of every entry --- one grid, two reading orders.}
\inv{This trivial-looking fact IS double counting. All of today's incidence arguments (B10, C6, D4) are this sentence applied to a cleverly chosen grid: rows = people, columns = committees, entries = memberships.}

%── A8 ──────────────────────────────────────────────────────────────────────────
\item Guarantee two integers differing by a multiple of 5?

\ans{$6$}
\method{Boxes are the residues mod 5; same residue $\Rightarrow$ difference divisible by 5.}
\inv{Residue classes are the most important pigeonholes in competition maths. Note the construction side: $1,2,3,4,5$ (one per residue) shows 5 integers do NOT suffice --- a guarantee question is only fully answered when both directions are shown.}

%── A9 ──────────────────────────────────────────────────────────────────────────
\item Sweets to guarantee 3 of one flavour (4 flavours)?

\ans{$4\cdot2+1=9$}
\method{Worst case: 2 of every flavour (8 sweets); the ninth forces a third somewhere.}
\inv{Strong pigeonhole: to force $m$ in one of $k$ boxes you need $k(m-1)+1$ pigeons. Re-derive A1 ($m=2$, $k=2$) and C1 from this one formula.}

%── A10 ─────────────────────────────────────────────────────────────────────────
\item Any 4 integers: two with even difference?

\ans{True}
\method{Two parity boxes, four integers --- some two share parity, and their difference is even.}
\inv{Sharper: among any 3 integers two share parity (A5). Among any 4 integers, must two differ by a multiple of 3? (Yes --- 3 boxes.) The box count, not the pigeon count, is the quantity to engineer.}

\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50\,s each)}}
%═══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}

%── B1 ──────────────────────────────────────────────────────────────────────────
\item 12 round-table seats: smallest number forcing an adjacent pair? \textit{\small(after TMUA 2019 P2 Q9)}

\ans{$7$}
\method{Six diners can sit in alternate seats (construction: seats $1,3,5,7,9,11$), so 6 does not force adjacency. With 7 diners, pair the seats into 6 adjacent couples $\{1,2\},\{3,4\},\ldots,\{11,12\}$: seven diners in six couples put two in the same couple --- adjacent. Answer: 7.}
\inv{Both halves matter: the construction shows 6 fails, the pigeonhole shows 7 works. TMUA multiple-choice hides this structure; BMO1 demands you write both. Generalise: $2n$ seats need $n+1$ diners.}

%── B2 ──────────────────────────────────────────────────────────────────────────
\item Guarantee two integers with difference divisible by 7?

\ans{$8$}
\method{Seven residue boxes mod 7.}
\inv{Follow-up with structure: among any 8 integers, two differ by a multiple of 7 --- but among any \emph{four} integers, must two SUM or differ to a multiple of 5? (Boxes $\{0\},\{1,4\},\{2,3\}$ mod 5: three boxes, four pigeons --- yes. Pairing residues that sum to 5 is the C2 trick in modular clothing.)}

%── B3 ──────────────────────────────────────────────────────────────────────────
\item Two of any 5 integers differ by a multiple of 4; five integers avoiding multiples-of-5 differences.

\ans{Pigeonhole on 4 residues; construction: $1,2,3,4,5$}
\method{Mod 4 there are only 4 residues, so 5 integers repeat one; equal residues difference $\equiv0$. For the second part, any complete residue system mod 5 works: $1,2,3,4,5$ have all differences $1$--$4$ mod 5, never $0$.}
\inv{Notice the numbers 4 and 5 trade places in the two halves --- the guarantee threshold for modulus $m$ is exactly $m+1$. State this as a single theorem covering A8, B2 and this question.}

%── B4 ──────────────────────────────────────────────────────────────────────────
\item Six from $\{1,\ldots,10\}$: two summing to 11? \textit{\small(after TMUA 2022 P2 Q8)}

\ans{Yes; the smallest forcing size is $6$}
\method{Pair the set into $\{1,10\},\{2,9\},\{3,8\},\{4,7\},\{5,6\}$: five boxes each summing to 11. Six chosen numbers land two in one box. Five numbers can dodge (take one from each pair, e.g.\ $1,2,3,4,5$).}
\inv{Sum-pairs are pigeonholes just as residues are. Design your own: from $\{1,\ldots,20\}$, how many numbers force two summing to 21? Two differing by 10? (Same boxes both times! Why?)}

%── B5 ──────────────────────────────────────────────────────────────────────────
\item Round-robin with 8 teams: matches and per-team games?

\ans{$\dbinom{8}{2}=28$ matches; each team plays $7$}
\method{A match is a pair of teams. Per team: everyone else once.}
\inv{Double-count check: $8\cdot7$ counts each match twice, $56/2=28$. \checkmark\ Real leagues double it (home and away: 56). How many matches in a league where each team plays each other $k$ times?}

%── B6 ──────────────────────────────────────────────────────────────────────────
\item Friendship counts total 84: how many friendships?

\ans{$42$}
\method{Handshake lemma: each friendship contributes 2 to the total.}
\inv{Could the counts have totalled 85? (No --- the degree sum is always even.) This parity fact silently solves impossible-configuration questions: e.g.\ can 7 people each have exactly 3 friends? ($7\cdot3=21$ odd --- impossible.)}

%── B7 ──────────────────────────────────────────────────────────────────────────
\item Four from $\{1,\ldots,6\}$: two consecutive?

\ans{Forced; largest safe choice has 3 elements, e.g.\ $\{1,3,5\}$}
\method{Boxes $\{1,2\},\{3,4\},\{5,6\}$: four numbers in three boxes put two in one box, and box-mates are consecutive. The construction $\{1,3,5\}$ (or $\{2,4,6\}$, $\{1,4,6\}$, \ldots) shows 3 can dodge.}
\inv{Connect back to sheet 2's D2: subsets of $\{1,\ldots,8\}$ with NO two consecutive were counted by Fibonacci (55 of them). Guarantee questions ask for the \emph{largest} such subset instead --- extremal vs enumerative, two faces of one constraint.}

%── B8 ──────────────────────────────────────────────────────────────────────────
\item Two party guests with equal handshake counts: why can't $0$ and $n-1$ coexist?

\ans{Someone shaking all $n-1$ hands has shaken the non-shaker's hand --- contradiction.}
\method{Possible counts are $0,1,\ldots,n-1$: that is $n$ boxes for $n$ people, so pigeonhole alone fails. But the extreme boxes are incompatible: a count of $n-1$ (shook everyone) and a count of $0$ (shook no one) cannot both occur. So at most $n-1$ counts are actually available for $n$ people --- pigeonhole now bites.}
\inv{This ``shrink the boxes by a contradiction'' move upgrades many pigeonhole arguments. D3 asks for the full formal write-up --- draft it now while the idea is warm.}

%── B9 ──────────────────────────────────────────────────────────────────────────
\item 25 sweets, 7 children: someone gets at least 4.

\ans{At least $\lceil25/7\rceil=4$}
\method{If every child had at most 3, the total would be at most $21<25$. Contradiction.}
\inv{The averaging form of pigeonhole: some value is $\geq$ the mean (and some is $\leq$). Convert to the strong form: $7\cdot3+1=22$ sweets already force a 4; we had 25. Which formulation writes up faster?}

%── B10 ─────────────────────────────────────────────────────────────────────────
\item 5 committees of 4; everyone on exactly 2: how many members?

\ans{$\dfrac{5\cdot4}{2}=10$}
\method{Count (person, committee) memberships two ways: committees give $5\cdot4=20$; people give $2\cdot(\text{members})$. So members $=10$.}
\inv{A7's grid made explicit: rows are people, columns committees. Whenever a puzzle says ``each X has $a$ Y's and each Y has $b$ X's'', the incidence count $a|X|=b|Y|$ falls out. D4 is this with one unknown.}

\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90\,s each)}}
%═══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}

%── C1 ──────────────────────────────────────────────────────────────────────────
\item 10 red, 10 green, 10 blue: guarantee 5 of some colour; guarantee 5 red?

\ans{Some colour: $3\cdot4+1=13$.\quad Five red: $20+4+1=25$.}
\method{\emph{Some colour} is plain strong pigeonhole: worst case 4 of each (12 beads), the 13th forces a fifth. \emph{Five red} is adversarial: the drawer hands you all 10 green and 10 blue first (20 useless beads), then 4 red --- so 24 can still fail, and the 25th guarantees a fifth red.}
\inv{The two answers differ because ``some colour'' lets any colour win, while ``red'' must survive the drawer burning every other colour first --- the A2 lesson (a pair vs a black pair) scaled up. What if you needed 5 red \emph{and} 5 green? Chase the worst case: burn blue first.}

%── C2 ──────────────────────────────────────────────────────────────────────────
\item Seven from $\{1,\ldots,12\}$: two sum to 13; six don't force.

\ans{Forced at 7; construction for 6: $\{1,2,3,4,5,6\}$}
\method{Boxes $\{1,12\},\{2,11\},\{3,10\},\{4,9\},\{5,8\},\{6,7\}$ --- six boxes each summing to 13. Seven numbers double up somewhere. Taking the smaller element of each pair ($1$--$6$) dodges with six numbers: all pair-sums are at most $11<13$.}
\inv{Verify the construction really works (largest sum $5+6=11$). Then a two-step hybrid: how many numbers from $\{1,\ldots,12\}$ force EITHER a pair summing to 13 OR a pair differing by 6? (The boxes coincide! Explain why $\{k,13-k\}$ and difference-6 pairs interlock.)}

%── C3 ──────────────────────────────────────────────────────────────────────────
\item Guarantee two integers with the same last digit?

\ans{$11$}
\method{Ten digit-boxes $0$--$9$.}
\inv{Same last digit $\iff$ difference divisible by 10 --- so this is A8/B2 with modulus 10. Chain it: among 11 integers, two differ by a multiple of 10; among how many, two differing by a multiple of 100? The modulus is the box count, always.}

%── C4 ──────────────────────────────────────────────────────────────────────────
\item Mutilated chessboard: 31 dominoes?

\ans{Impossible.}
\method{Colour the board alternately. Opposite corners share a colour, so the mutilated board has 32 squares of one colour and 30 of the other. Every domino covers one square of each colour, so 31 dominoes cover 31 of each --- but $31\neq32$. No tiling exists. $\square$}
\inv{The first \emph{invariant} argument of the pillar: a quantity every move preserves (colour balance) separates reachable from unreachable. Sheet 7 builds a whole section on this. Try: remove two squares of OPPOSITE colours anywhere --- can the remaining 62 always be tiled? (Yes --- Gomory's beautiful cycle argument; look it up after attempting.)}

%── C5 ──────────────────────────────────────────────────────────────────────────
\item Group size to guarantee two share both birth month and weekday?

\ans{$12\cdot7+1=85$}
\method{The pigeonholes are (month, weekday) pairs: $12\times7=84$ boxes. With 85 people, two land in the same box --- same month AND same weekday.}
\inv{Two independent attributes multiply their box counts, they do not add --- $84$, not $19$. This product-of-dimensions idea reappears whenever you pigeonhole on several coordinates at once (residues mod different primes, grid cells). How many boxes if you also track blood type (4 kinds)?}

%── C6 ──────────────────────────────────────────────────────────────────────────
\item 20 clubs $\times$ 6 students; each student in 3 clubs: student count?

\ans{$\dfrac{20\cdot6}{3}=40$}
\method{Memberships counted by clubs: 120. Counted by students: $3S$. So $S=40$.}
\inv{Add a consistency question: could the same school run 21 clubs of 6 with everyone in exactly 3? ($126/3=42$ --- fine). And 20 clubs of 7? ($140/3$ is not an integer --- impossible!). Divisibility of the double count is a hidden feasibility test.}

%── C7 ──────────────────────────────────────────────────────────────────────────
\item $1$--$9$ around a circle: three adjacent summing $\geq15$.

\ans{Some three adjacent numbers sum to at least 15.}
\method{The nine positions split into three \emph{disjoint} blocks of three consecutive seats. The three block-sums total $1+2+\cdots+9=45$, so some block sums to at least $\lceil45/3\rceil=15$. $\square$}
\inv{Averaging pigeonhole on engineered boxes. Sharper (harder): show some three adjacent sum $\geq16$ fails --- find an arrangement whose every adjacent triple sums $\leq16$ but... actually can every triple be $\leq15$? The nine \emph{overlapping} triples sum to $3\cdot45=135=9\cdot15$ exactly --- so all nine triples equal 15, forcing impossible structure. Chase this to decide whether 15 is tight.}

%── C8 ──────────────────────────────────────────────────────────────────────────
\item Five from $\{1,\ldots,8\}$: two differing by 4.

\ans{Forced.}
\method{Boxes $\{1,5\},\{2,6\},\{3,7\},\{4,8\}$: four boxes whose mates differ by exactly 4. Five numbers put two in one box. $\square$}
\inv{Largest safe set: 4 (one per box --- e.g.\ $\{1,2,3,4\}$). Variant with difference 3 on $\{1,\ldots,9\}$: boxes $\{1,4\},\{2,5\},\{3,6\},\{7,?\}$... the pairing frays. Build correct boxes for difference 3 (chains $1\to4\to7$, $2\to5\to8$, $3\to6\to9$: pick alternate links) and find the true threshold.}

\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(SMC / BMO1 difficulty)}}
%═══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}

%── D1 ──────────────────────────────────────────────────────────────────────────
\item Five points in a unit square: two within $\frac{\sqrt2}{2}$.

\ans{Proof below.}
\method{\emph{Proof.} Divide the unit square into four closed $\frac12\times\frac12$ quarter-squares. Five points in four quarters: by the pigeonhole principle two points, say $P$ and $Q$, lie in the same quarter. Any two points of a $\frac12\times\frac12$ square are at distance at most its diagonal, $\frac{\sqrt2}{2}$. Hence $|PQ|\leq\frac{\sqrt2}{2}$. $\square$}
\inv{The write-up has three named steps: build boxes, apply pigeonhole, bound within a box --- every geometric pigeonhole proof has this skeleton. Is $\frac{\sqrt2}{2}$ best possible for five points? (Place four corners + centre and measure.) What about \emph{four} points --- what distance can they all keep?}

%── D2 ──────────────────────────────────────────────────────────────────────────
\item Five lattice points: some midpoint is a lattice point.

\ans{Proof below.}
\method{\emph{Proof.} Classify each point $(x,y)$ by the parities of its coordinates: (even, even), (even, odd), (odd, even), (odd, odd) --- four classes. Five points give two in the same class, say $P=(a,b)$ and $Q=(c,d)$ with $a\equiv c$ and $b\equiv d\pmod 2$. Then $a+c$ and $b+d$ are both even, so the midpoint $\left(\frac{a+c}{2},\frac{b+d}{2}\right)$ has integer coordinates. $\square$}
\inv{Four points can dodge: $(0,0),(0,1),(1,0),(1,1)$ --- one per parity class; check all six midpoints fail. So 5 is exact. Generalise to 3D: how many lattice points force a lattice midpoint? ($2^3+1=9$.) And what forces a lattice \emph{centroid} of some three points? (Much subtler --- sit with it.)}

%── D3 ──────────────────────────────────────────────────────────────────────────
\item The number of people with an odd handshake count is even.

\ans{Proof below.}
\method{\emph{Proof.} Let the guests be the vertices of a graph, with an edge for each handshake; write $d(v)$ for the number of hands $v$ shook. Every handshake has two participants, so summing over all guests counts each handshake twice: $\sum_{v}d(v)=2H$, where $H$ is the total number of handshakes --- an even number. Split the sum by parity: $\sum_{d(v)\text{ even}}d(v)$ is even (a sum of even numbers), and the whole sum is even, so $\sum_{d(v)\text{ odd}}d(v)$ is even too. But that last sum is a sum of odd numbers, and a sum of odd numbers is even only when there is an even count of them. Hence the number of guests with an odd handshake count is even. $\square$}
\inv{This is the Handshake Lemma's famous corollary. It instantly kills impossible configurations: can 9 people each shake exactly 3 hands? (Nine odds is odd --- no.) Sit with the parity step ``a sum of odd numbers is even iff there is an even number of them'' until it is obvious; it is the whole proof.}

%── D4 ──────────────────────────────────────────────────────────────────────────
\item 20 students, each solving exactly 2 of 5 problems: two share a problem-pair.

\ans{Proof below; some pair chosen by $\geq2$ students (in fact $\geq\lceil20/10\rceil=2$).}
\method{\emph{Proof.} Each student's choice of two problems from five is an unordered pair, and there are exactly $\binom{5}{2}=10$ such pairs. Regard the 10 pairs as pigeonholes and the 20 students as pigeons. Since $20>10$, by the pigeonhole principle some pigeonhole holds at least two students --- that is, some pair of problems was chosen by at least two students, who therefore solved exactly the same pair. (Strong pigeonhole sharpens this: at least $\lceil20/10\rceil=2$ share the most popular pair, and averaging shows some pair is chosen by at least 2 regardless of the distribution.) $\square$}
\inv{Naming the pigeonholes as the $\binom{5}{2}$ pairs --- rather than the problems --- is the whole idea; a student who lists the wrong boxes gets nothing. Sharpen it: with 21 students must some pair be chosen three times? (No --- $21<2\cdot10+1$.) Find the exact threshold that forces a triple.}

%── D5 ──────────────────────────────────────────────────────────────────────────
\item Six people: three mutual friends or three mutual strangers.

\ans{Proof below.}
\method{\emph{Proof.} Fix a person $X$. The other five people split into friends of $X$ and strangers to $X$; by pigeonhole one class contains at least $\lceil5/2\rceil=3$ people --- say (without loss of generality) $A,B,C$ are all friends of $X$. Now examine the three relations among $A,B,C$: if any pair, say $A$ and $B$, are friends, then $X,A,B$ are three mutual friends. Otherwise all three pairs among $A,B,C$ are strangers, and $A,B,C$ are three mutual strangers. Either way the claim holds. (If the majority class was strangers, the same argument runs with roles swapped.) $\square$}
\inv{This is the Ramsey number statement $R(3,3)\leq6$. Show 5 people are not enough: arrange five people in a cycle where neighbours are friends and non-neighbours strangers, and verify no monochromatic triangle. Hence $R(3,3)=6$ exactly. The write-up's ``without loss of generality'' is load-bearing --- say why the symmetry justifies it.}

\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Top 5 Patterns Today}
%═══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}[leftmargin=2em, itemsep=4pt]
  \item \textbf{Basic pigeonhole.} $k+1$ objects in $k$ boxes: some box holds two. The craft is in \emph{designing} the boxes: residues, sum-pairs, quarter-squares, parity classes.
  \item \textbf{Strong pigeonhole.} Forcing $m$ alike among $k$ types needs $k(m-1)+1$ objects. Equivalently: some box holds at least the ceiling of the average.
  \item \textbf{Worst case first.} Every guarantee number is (largest configuration that dodges) $+1$. A complete answer exhibits the dodge AND proves the force.
  \item \textbf{Double counting.} Count (X, Y) incidences by X and by Y: $a|X|=b|Y|$. Grids, committees, degree sums --- and an integrality feasibility test for free.
  \item \textbf{Shrink the boxes by contradiction.} When boxes equal pigeons, show two boxes cannot both be occupied (counts 0 and $n-1$) --- then pigeonhole closes the argument.
\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Common Traps to Avoid}
%═══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}[leftmargin=2em, itemsep=4pt]
  \item \textbf{Answering the force without the dodge.} ``7 diners force adjacency'' is half marks until you seat 6 who don't. Guarantee $=$ construction $+$ pigeonhole.
  \item \textbf{Off-by-one in strong pigeonhole.} It is $k(m-1)+1$, not $km$ or $km+1$. Re-derive from the worst case each time rather than recalling.
  \item \textbf{Boxes that aren't disjoint or don't cover.} Sum-pair and difference boxes must partition the set --- check both properties explicitly before counting pigeons.
  \item \textbf{Double counting with one side wrong.} ``Each of 20 clubs of 7, everyone in 3'' has no integer solution --- test divisibility before trusting a puzzle's premises.
  \item \textbf{Vague write-ups.} ``By pigeonhole, done'' scores nothing. Name the pigeons, name the boxes, state which two collide and what that collision gives you.
\end{enumerate}

\SpeedClosing{``The pigeonhole principle proves existence without ever finding the thing ---
mathematics at its most confident.''}

\end{document}
