\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Combinatorics}{3}

\begin{document}

\SpeedTitleBlock{Daily Combinatorics Drill \#3 --- Answers \& Investigations}
\noindent\textit{New toolkit today: Blocks \& Gaps $\cdot$ Circular Arrangements $\cdot$ Repeated Letters $\cdot$ Forced-Order Tricks.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15\,s each)}}
%═══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}

%── A1 ──────────────────────────────────────────────────────────────────────────
\item In how many ways can 5 people stand in a row?

\ans{$5!=120$}
\method{Distinct objects in a row: factorial. Instant.}
\inv{Sheet 1's opening question returns as pure recall --- that is the point of Section A. If this took more than three seconds, redo sheet 1's A section tomorrow.}

%── A2 ──────────────────────────────────────────────────────────────────────────
\item How many arrangements of the letters of LEMON?

\ans{$5!=120$}
\method{Five \emph{distinct} letters --- same question as A1 in costume.}
\inv{Spot the isomorphism reflex: people in rows, distinct letters, distinct books are the same problem. The next question breaks the spell --- watch what changes.}

%── A3 ──────────────────────────────────────────────────────────────────────────
\item How many arrangements of the letters of BANANA?

\ans{$\dfrac{6!}{3!\,2!}=60$}
\method{Six letters: A$\times$3, N$\times$2, B$\times$1. Divide $6!$ by the repeats: $\frac{720}{6\cdot2}$.}
\inv{Why division works: the 3 A's can be permuted in $3!$ ways without changing the word, so each distinct word was counted exactly $3!\,2!$ times among the $6!$. This ``quotient out the symmetry'' idea is the same one that will give circular arrangements in A4.}

%── A4 ──────────────────────────────────────────────────────────────────────────
\item Four people around a round table (rotations the same)?

\ans{$(4-1)!=6$}
\method{Fix one person to kill the rotational symmetry; arrange the remaining 3 freely.}
\inv{Alternative view: $\frac{4!}{4}$ --- each circular arrangement is counted 4 times among row arrangements, once per rotation. Check both give 6. When would you also divide by 2? (When the table can be ``flipped'' --- necklaces, handshake circles.)}

%── A5 ──────────────────────────────────────────────────────────────────────────
\item Six in a row, Ali first and Zara last?

\ans{$4!=24$}
\method{Two positions forced, four people free.}
\inv{Forced positions contribute factor 1 --- cousins of sheet 2's forced aces. Variant: Ali \emph{or} Zara first (and no other constraint): $2\cdot5!=240$. Why can you not just double 24?}

%── A6 ──────────────────────────────────────────────────────────────────────────
\item Three identical red flags and two identical blue flags in a row?

\ans{$\dbinom{5}{2}=10$}
\method{Only the \emph{positions} of the blue flags matter: choose 2 places from 5. Equivalently $\frac{5!}{3!2!}$.}
\inv{Repeated-letter arrangements and binomial coefficients are the same formula meeting in the middle: $\frac{5!}{3!2!}=\binom{5}{2}$. Every binary string question (sheet 2, B6) is a flag question. What is the three-colour version? (The multinomial coefficient --- sheet 4.)}

%── A7 ──────────────────────────────────────────────────────────────────────────
\item Arrangements of NOON?

\ans{$\dfrac{4!}{2!\,2!}=6$}
\method{Four letters, two pairs: $\frac{24}{4}$.}
\inv{List all six to feel the formula: NOON, NONO, NNOO, OONN, ONON, ONNO. Any word whose repeat signature is $(2,2)$ has 6 arrangements --- signature, not spelling, decides the count.}

%── A8 ──────────────────────────────────────────────────────────────────────────
\item Six people around a round table (rotations the same)?

\ans{$(6-1)!=120$}
\method{Fix one person, arrange 5.}
\inv{A dinner-party host claims seating 6 guests differently every night is possible for four months. Check: 120 nights $\approx$ four months. For how long could 8 guests last? (Nearly 14 years.)}

%── A9 ──────────────────────────────────────────────────────────────────────────
\item True or false: 8 people round a table $=$ 7 people in a row?

\ans{True --- both are $7!=5040$}
\method{Circular $n$ $=(n-1)!$ $=$ row arrangements of $n-1$. Fixing one diner \emph{is} reducing to a row of the rest.}
\inv{TMUA-style equivalence spotting. Decide true/false: ``$2n$ people in a circle $=$ $n$ people in a row, for some $n>1$.'' ($(2n-1)!=n!$ forces $2n-1=n$, impossible for $n>1$ --- false.)}

%── A10 ─────────────────────────────────────────────────────────────────────────
\item Arrangements of TATTY?

\ans{$\dfrac{5!}{3!}=20$}
\method{Five letters with T$\times$3: divide by $3!$.}
\inv{Rank the words BANANA (60), TATTY (20), LEVEL (30), LEMON (120) by arrangement count \emph{before} computing, using only the repeat structure. Fluency at reading the ``signature'' $(3,2,1)$ vs $(3,1,1)$ vs $(2,2,1)$ is the skill.}

\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50\,s each)}}
%═══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}

%── B1 ──────────────────────────────────────────────────────────────────────────
\item Six in a row, Priya and Quinn together?

\ans{$2\cdot5!=240$}
\method{Glue the pair (block has 2 internal orders), arrange 5 objects.}
\inv{Sheet 1 met this in Section C; now it is a B-level reflex --- the spiral is deliberate. Speed target: write ``$2\cdot5!$'' within five seconds of finishing reading.}

%── B2 ──────────────────────────────────────────────────────────────────────────
\item Six in a row, Priya and Quinn \emph{not} together?

\ans{$720-240=480$}
\method{Complement of B1.}
\inv{Reprove via gaps: seat the other 4 ($4!$), then place P and Q in different gaps of the 5 available: $4!\cdot5\cdot4=480$. The gap method scales to ``no two of \emph{three} people adjacent'' where complements get messy --- see D5.}

%── B3 ──────────────────────────────────────────────────────────────────────────
\item Arrangements of MISSISSIPPI?

\ans{$\dfrac{11!}{4!\,4!\,2!}=34\,650$}
\method{11 letters: I$\times$4, S$\times$4, P$\times$2, M$\times$1. One division, no listing.}
\inv{Compute it calculator-free the clean way: $\frac{11!}{4!4!2!}=\binom{11}{4}\binom{7}{4}\binom{3}{2}=330\cdot35\cdot3$. Choosing positions letter-by-letter both proves the formula and organises the arithmetic.}

%── B4 ──────────────────────────────────────────────────────────────────────────
\item Five boys, three girls in a row, no two girls adjacent?

\ans{$5!\cdot(6\cdot5\cdot4)=120\cdot120=14\,400$}
\method{Gap method: arrange the boys ($5!$), creating 6 gaps (including the two ends); place the three distinct girls in three \emph{different} gaps, in order: $6\cdot5\cdot4$.}
\inv{Why does gluing/complement fail here? ``No two of three adjacent'' has too many overlapping bad events. The gap method sidesteps inclusion--exclusion entirely --- but inclusion--exclusion (sheet 5) will handle what gaps cannot.}

%── B5 ──────────────────────────────────────────────────────────────────────────
\item Four men and four women in a row, alternating?

\ans{$2\cdot4!\cdot4!=1\,152$}
\method{Two gender patterns (MWMW\ldots{} or WMWM\ldots), then the men fill their 4 slots in $4!$ ways and the women theirs in $4!$.}
\inv{With 5 men and 4 women only one pattern survives (MWMWMWMWM): $5!\cdot4!=2880$, no factor 2. The factor-of-2 question is always: does the \emph{pattern} have a choice, or is it forced?}

%── B6 ──────────────────────────────────────────────────────────────────────────
\item Six around a round table, two particular people together?

\ans{$2\cdot(5-1)!=48$}
\method{Glue the pair into a block: 5 objects round a table $=(5-1)!$, times 2 internal orders.}
\inv{Row version was $2\cdot5!=240$; circle version is $2\cdot4!=48$. The ratio is 5 --- exactly the number of rotations quotiented out. Check this ratio logic predicts the answer for 8 people with a glued pair.}

%── B7 ──────────────────────────────────────────────────────────────────────────
\item Six-digit strings from $1,1,2,2,3,3$?

\ans{$\dfrac{6!}{2!\,2!\,2!}=90$}
\method{Three pairs: divide $720$ by $2^3$.}
\inv{Of these 90, how many are palindromes? (A palindrome is fixed by its first three digits, which must use each of 1, 2, 3 exactly once: $3!=6$.) Symmetry constraints slash repeated-letter counts --- try ``how many read the same upside down'' style variants.}

%── B8 ──────────────────────────────────────────────────────────────────────────
\item Arrangements of LEVEL?

\ans{$\dfrac{5!}{2!\,2!}=30$}
\method{L$\times$2, E$\times$2, V: divide $120$ by 4.}
\inv{How many of the 30 are palindromes? (Middle letter forced to V; first two letters L, E in some order determine the rest: 2.) Notice palindrome counting always halves the alphabet workload: only half the word is free.}

%── B9 ──────────────────────────────────────────────────────────────────────────
\item Seven books, trilogy in volume order (not necessarily adjacent)?

\ans{$\dfrac{7!}{3!}=840$}
\method{Forced-order trick: of the $3!$ orderings of the trilogy within any arrangement, exactly one is correct --- so divide $7!$ by $3!$. No gluing: they need not be adjacent.}
\inv{Compare the three trilogy conditions: adjacent-any-order ($3!\cdot5!=720$), adjacent-in-order ($5!=120$), in-order-anywhere ($840$). Rank them by restrictiveness \emph{before} computing next time. The divide-by-order trick is sheet 1's ``strictly increasing digits'' insight, now industrialised.}

%── B10 ─────────────────────────────────────────────────────────────────────────
\item Round-table seatings of $n$ people $=720$. Find $n$.

\ans{$n=7$}
\method{$(n-1)!=720=6!$, so $n=7$.}
\inv{Factorials are so sparse that ``$=720$'' identifies $6!$ instantly --- recall sheet 1's ladder. Solve also $(n-1)!\cdot2=240$ (a glued-pair circle count): $n=6$.}

\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90\,s each)}}
%═══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}

%── C1 ──────────────────────────────────────────────────────────────────────────
\item Eight in a row: three friends consecutive, two others (non-friends) not adjacent?

\ans{$3!\cdot6!-3!\cdot2!\cdot5!=4\,320-1\,440=2\,880$}
\method{Layer the constraints (sheet's D2 pattern, arriving early). Glue the trio: $3!\cdot6!=4\,320$ arrangements. From these subtract those where the two enemies are also adjacent --- glue them too: $3!\cdot2!\cdot5!=1\,440$. Answer $2\,880$.}
\inv{Generalise the single-constraint count: $k$ specified people consecutive among $n$ gives $k!\,(n-k+1)!$ --- check it against B1. Then redo this question with the gap method instead of the complement and confirm agreement.}

%── C2 ──────────────────────────────────────────────────────────────────────────
\item Same eight, no refusals, friends consecutive \emph{and} in height order?

\ans{$6!=720$}
\method{The block's internal order is now forced: glue with factor 1, then 6 objects in a row.}
\inv{``Consecutive AND ordered'' = glue with factor 1; ``ordered but scattered'' = divide by $3!$ (B9). Keep the two tricks distinct in your head --- mixing them up is a classic exam slip.}

%── C3 ──────────────────────────────────────────────────────────────────────────
\item Arrangements of DIVIDED without all three D's together?

\ans{$\dfrac{7!}{3!\,2!}-\dfrac{5!}{2!}=420-60=360$}
\method{Total arrangements: $\frac{7!}{3!2!}=420$. All-D's-together: glue [DDD] (identical letters, factor 1) and arrange it with I, I, V, E: $\frac{5!}{2!}=60$. Subtract.}
\inv{Careful distinction: the complement of ``all three together'' still allows \emph{two} D's adjacent. Count the arrangements where no two D's are adjacent at all (gap method on I, I, V, E: $\frac{4!}{2!}\cdot\binom{5}{3}=120$) and check $120<360$ makes sense.}

%── C4 ──────────────────────────────────────────────────────────────────────────
\item Four men, four women alternating around a round table?

\ans{$3!\cdot4!=144$}
\method{Seat the men first: circular, so $(4-1)!=6$. Their seating fixes 4 alternating gaps, into which the women go in $4!$ ways. (No extra factor 2: in a circle the two ``patterns'' are rotations of each other.)}
\inv{Row version (B5) had the pattern factor 2; the circle ate it. Trace exactly where: rotating an MWMW\ldots{} row by one seat turns it into WMWM\ldots{} --- rotations already identify the two patterns. Symmetries must be counted once, not twice.}

%── C5 ──────────────────────────────────────────────────────────────────────────
\item Arrangements of CIRCLE with the two C's apart?

\ans{$\dfrac{6!}{2!}-5!=360-120=240$}
\method{Total (with C repeated): 360. Together: glue CC (identical, no internal factor): $5!=120$. Subtract.}
\inv{Redo via gaps: arrange I, R, L, E ($4!$), choose 2 of the 5 gaps for the identical C's ($\binom{5}{2}$): $24\cdot10=240$. \checkmark\ Complement and gaps agree --- and for \emph{identical} letters the gap count uses $\binom{}{}$, not a falling product. Spot why.}

%── C6 ──────────────────────────────────────────────────────────────────────────
\item Five-digit even numbers over $40\,000$ using $1,2,3,4,5$ once each?

\ans{$18$}
\method{Two interacting constraints (units even, leading digit 4 or 5) --- case on the units digit, because it consumes a candidate leading digit or not. \emph{Units 2:} leading digit from $\{4,5\}$, rest free: $2\cdot3!=12$. \emph{Units 4:} leading digit must be 5, rest free: $3!=6$. Total $18$.}
\inv{Why case on units rather than the leading digit? Try it the other way and watch the option-counts stop being uniform (sheet 1's C1 lesson). Extension: how many are divisible by 4? (Endings 12, 24, 32, 52; now impose the $>40\,000$ condition on each case.)}

%── C7 ──────────────────────────────────────────────────────────────────────────
\item Three maths, two physics, two chemistry books, subjects together?

\ans{$3!\cdot(3!\cdot2!\cdot2!)=6\cdot24=144$}
\method{Arrange the three subject blocks ($3!$), then the books inside each block ($3!,2!,2!$).}
\inv{Blocks-of-blocks compose: shelves of subjects of books of chapters\ldots{} each level multiplies independently. Now add the constraint ``maths must be left of chemistry'' --- which factor changes, and to what? (The $3!$ becomes 3: divide by the $2!$ orderings of the M--C pair.)}

%── C8 ──────────────────────────────────────────────────────────────────────────
\item Eight around a round table, two people refusing to be adjacent?

\ans{$7!-2\cdot6!=5\,040-1\,440=3\,600$}
\method{All seatings ($7!$) minus glued-pair seatings ($2\cdot6!$).}
\inv{Gap version: seat the 6 others round the table ($5!$), then put the two enemies into 2 of the 6 gaps: $5!\cdot6\cdot5=3600$. \checkmark\ Both routes again --- build the habit of solving restricted seatings twice until they always agree first time.}

\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(SMC difficulty)}}
%═══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}

%── D1 ──────────────────────────────────────────────────────────────────────────
\item Seven-digit numbers from $1$--$7$: odds increasing, evens decreasing?

\ans{$\dbinom{7}{4}=35$}
\method{Choose which 4 of the 7 positions hold the odd digits: $\binom{7}{4}=35$. Everything else is forced --- the odds $1,3,5,7$ must fill their positions in increasing order (one way), and the evens $6,4,2$ fill theirs in decreasing order (one way). A count that looks like it needs $7!$ cases collapses to a single choice.}
\inv{Equivalently by double forced-order division: $\frac{7!}{4!\,3!}=35$ --- divide by the orderings of the odds AND of the evens. With only the odds constrained the count is $\frac{7!}{4!}=210$; check both against the choose-positions view. This ``order is forced, so only positions matter'' insight is the SMC's favourite disguise for $\binom{n}{k}$.}

%── D2 ──────────────────────────────────────────────────────────────────────────
\item Ten children in a circle: Mia--Noah adjacent, Omar--Priya not?

\ans{$2\cdot8!-4\cdot7!=80\,640-20\,160=60\,480$}
\method{Layer the constraints. Glue Mia--Noah: $2\cdot(9-1)!=80\,640$ circles. From these subtract the ones where Omar--Priya are also adjacent (glue both pairs): $2\cdot2\cdot(8-1)!=20\,160$. Answer: $60\,480$.}
\inv{The pattern ``force the easy constraint, complement the hard one'' handles most two-constraint arrangement problems. Try three pairs: A--B adjacent, C--D adjacent, E--F \emph{not} adjacent, 10 people in a circle. (Same layering, one more glue.)}

%── D3 ──────────────────────────────────────────────────────────────────────────
\item Arrangements of AABBCC with no two identical letters adjacent?

\ans{$30$}
\method{Inclusion--exclusion on the ``bad pair'' events, computed cleanly by gluing. Total: $\frac{6!}{2!2!2!}=90$. Glue AA: $\frac{5!}{2!2!}=30$, same for BB, CC. Glue two pairs: $\frac{4!}{2!}=12$ each (3 ways). Glue all three: $3!=6$. Count with no glued pair $=90-3\cdot30+3\cdot12-6=30$.}
\inv{This is inclusion--exclusion arriving a sheet early --- sheet 5 makes it systematic. For now, verify the answer differently: by symmetry all $3!$ letter-labelings are equivalent, so just count patterns like ABCABC, ABCACB, \ldots{} and multiply. How many \emph{patterns} are there? ($30/3!=5$.)}

%── D4 ──────────────────────────────────────────────────────────────────────────
\item Eight in a row, exactly two of the three sisters adjacent?

\ans{$8!-14\,400-4\,320=21\,600$}
\method{Three disjoint outcomes partition all $8!=40\,320$ arrangements: \emph{no} adjacent sister-pair, \emph{exactly one} adjacent sister-pair, and \emph{two} adjacent sister-pairs (which is precisely ``all three consecutive''). \emph{None adjacent} (gap method): $5!\cdot6\cdot5\cdot4=14\,400$. \emph{All three consecutive}: $3!\cdot6!=4\,320$. ``Exactly two sisters adjacent'' is the middle category, so it equals $40\,320-14\,400-4\,320=21\,600$.}
\inv{Check the borderline case that makes this work: in a block of three sisters, the outer two are NOT adjacent to each other --- adjacency is about neighbouring seats, not membership of a block. Redo the problem for ``at least two adjacent'' ($40\,320-14\,400=25\,920$) and confirm the three-way split is consistent.}

%── D5 ──────────────────────────────────────────────────────────────────────────
\item Eight knights round a table: three feuders pairwise non-adjacent, King and Queen together?

\ans{$3!\cdot2\cdot(4\cdot3\cdot2)=288$}
\method{Layer: glue first, then gaps. Glue the King--Queen pair (2 internal orders); the peaceful side is now 4 units (the KQ block plus 3 other knights). Seat those 4 units round the table: $(4-1)!=6$ circular orders, creating exactly 4 gaps. Place the three feuding knights into three different gaps (one each --- sharing a gap would make two feuders adjacent): $4\cdot3\cdot2=24$. Total $6\cdot2\cdot24=288$.}
\inv{Check the order of operations matters: try placing the feuders first and gluing second, and watch the gap count become ill-defined --- constraints that \emph{create} structure (glue) must precede constraints that \emph{consume} it (gaps). Then generalise the no-royals version: $n$ knights, $k$ feuding, count $=(n-k-1)!\cdot\frac{(n-k)!}{(n-2k)!}$; test it at $n=8$, $k=3$ (1440) against a hand count of a small case like $n=5$, $k=2$.}

\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Top 5 Patterns Today}
%═══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}[leftmargin=2em, itemsep=4pt]
  \item \textbf{Glue blocks.} ``Must be together'' $\Rightarrow$ treat as one object, multiply by internal orders (which is 1 if the glued items are identical or their order is forced).
  \item \textbf{Count gaps.} ``No two adjacent'' $\Rightarrow$ seat the unrestricted objects, then place the restricted ones into distinct gaps. Rows of $m$ give $m+1$ gaps; circles of $m$ give $m$ gaps.
  \item \textbf{Divide by forced order.} If $k$ objects must appear in one specified relative order (not necessarily adjacent), count freely and divide by $k!$.
  \item \textbf{Fix a seat in circles.} Rotational symmetry costs one factor of $n$: circular $n$ $=(n-1)!$. Ask separately whether reflections are also ``the same''.
  \item \textbf{Layer constraints.} Force the ``must'' constraints by gluing first, then handle the ``must not'' by complement or gaps inside that restricted world.
\end{enumerate}

%═══════════════════════════════════════════════════════════════════════════════
\section*{Common Traps to Avoid}
%═══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}[leftmargin=2em, itemsep=4pt]
  \item \textbf{Multiplying by internal orders of identical letters.} Gluing the two E's of SPEED is a factor 1, not 2. Only \emph{distinguishable} contents of a block get a factorial.
  \item \textbf{Row gaps vs circle gaps.} Five people create 6 gaps in a row but only 5 in a circle. Using the row count at a round table silently overcounts.
  \item \textbf{Double-counting the alternating patterns in a circle.} MWMW\ldots{} and WMWM\ldots{} are rotations of each other at a round table --- the row-version factor 2 must be dropped (C4).
  \item \textbf{Confusing ``in order'' with ``adjacent''.} Volume order scattered along the shelf divides by $3!$; volumes boxed together glues a block. Different tricks, different answers (B9 vs C1/C2).
  \item \textbf{Treating ``exactly one adjacent pair'' as total minus none-adjacent.} That difference counts \emph{at least one} pair. The all-together case must be carved out separately (D4).
\end{enumerate}

\SpeedClosing{``Glue what must be together, seat what must be apart last ---
the gaps will tell you how much room rebellion really has.''}

\end{document}
