\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Combinatorics}{6}

\begin{document}

\SpeedTitleBlock{Daily Combinatorics Drill \#6}
\noindent\textit{New toolkit today: Pigeonhole Principle (basic \& strong) $\cdot$ Worst-Case Thinking $\cdot$ Double Counting $\cdot$ The Handshake Lemma.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Guarantees on sight. Ask: what is the worst case?}

\begin{enumerate}

\item A drawer holds 10 black and 10 white socks. How many must be drawn (in the dark) to guarantee a matching pair?

\item Same drawer: how many draws guarantee a \emph{black} pair?

\item True or false: among any 13 people, two share a birth month.

\item How many people are needed to guarantee two were born on the same day of the week?

\item How many integers must you take to guarantee two of the same parity?

\item Six friends each count their handshakes at a party; the counts total 18. How many handshakes happened?

\item True or false: in any grid of numbers, the sum of the row totals equals the sum of the column totals.

\item How many integers must you take to guarantee two whose difference is a multiple of 5?

\item A bag has sweets of 4 flavours. How many sweets guarantee 3 of one flavour?

\item True or false: among any 4 integers, some two have an even difference.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Name the pigeonholes, or name what is being counted twice.}

\begin{enumerate}

\item A round table has 12 seats. What is the smallest number of diners that forces two of them to sit in adjacent seats? \textit{\small(after TMUA 2019 P2 Q9)}

\item How many integers must you take to guarantee two whose difference is divisible by 7?

\item Show that among any 5 integers, two have a difference divisible by 4. Then write down 5 integers of which no two differ by a multiple of \emph{5}.

\item Six numbers are chosen from $\{1,2,\ldots,10\}$. Must two of them sum to 11? What is the smallest choice size that forces this? \textit{\small(after TMUA 2022 P2 Q8)}

\item Eight teams play a round-robin (each pair meets once). How many matches are played, and how many does each team play?

\item In a class of 30, each student writes down their number of friends within the class, and the numbers total 84. How many friendships are there?

\item Four numbers are chosen from $\{1,2,\ldots,6\}$. Show that two of them must be consecutive, and exhibit a largest possible choice with no two consecutive.

\item At any party with at least two people, two attendees have shaken the same number of hands. What makes the counts $0$ and $n-1$ incompatible?

\item Twenty-five sweets are shared among 7 children. Show that some child receives at least 4.

\item A society has 5 committees, each with 4 members, and every member sits on exactly 2 committees. How many members does the society have?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Strong pigeonhole and structural double counts. Build the boxes before you count the pigeons.}

\begin{enumerate}

\item A jar holds beads of 3 colours. How many beads guarantee 5 of one colour?

\item Seven distinct numbers are chosen from $\{1,2,\ldots,12\}$. Show that two of them sum to 13, and show that six numbers are not enough to force this.

\item How many integers must you take to guarantee two ending in the same digit?

\item An $8\times8$ chessboard has two diagonally opposite corner squares removed. Can the remaining 62 squares be tiled by 31 dominoes? Justify your answer.

\item How large must a group be to guarantee that 4 of its members share a birth month?

\item A school runs 20 clubs, each with exactly 6 students, and every student belongs to exactly 3 clubs. How many students does the school have?

\item The numbers $1,2,\ldots,9$ are arranged around a circle in any order. Show that some three adjacent numbers have sum at least 15.

\item Five distinct numbers are chosen from $\{1,2,\ldots,8\}$. Show that two of them differ by exactly 4.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(SMC / BMO1 difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Determine, with proof. Model your write-up on the answer file's --- the argument is the answer.}

\begin{enumerate}

\item Five points lie in a unit square. Prove that two of them are at distance at most $\dfrac{\sqrt2}{2}$.

\item Five points have integer coordinates. Prove that the midpoint of some pair of them also has integer coordinates.

\item At a party with $n\geq2$ people, prove that two people have shaken hands with exactly the same number of fellow guests.

\item Twenty students each solve exactly two of the five problems on a quiz, and every problem is solved by the same number $k$ of students. Determine $k$, with proof.

\item Among any six people, any two of whom are either friends or strangers, prove that there exist three mutual friends or three mutual strangers.

\end{enumerate}

\SpeedClosing{``The pigeonhole principle proves existence without ever finding the thing ---
mathematics at its most confident.''}

\end{document}
