\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Combinatorics}{3}

\begin{document}

\SpeedTitleBlock{Daily Combinatorics Drill \#3}
\noindent\textit{New toolkit today: Blocks \& Gaps $\cdot$ Circular Arrangements $\cdot$ Repeated Letters $\cdot$ Forced-Order Tricks.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Rows, circles, repeated letters. One formula each. Pure recognition.}

\begin{enumerate}

\item In how many ways can 5 people stand in a row?

\item How many arrangements are there of the letters of LEMON?

\item How many arrangements are there of the letters of BANANA?

\item In how many ways can 4 people sit around a round table? (Rotations count as the same.)

\item Six people stand in a row with Ali first and Zara last. How many arrangements?

\item How many ways can 3 identical red flags and 2 identical blue flags be arranged in a row?

\item How many arrangements of the letters of SPEED have the two E's together?

\item In how many ways can 6 people sit around a round table? (Rotations count as the same.)

\item True or false: the number of ways to seat 8 people around a round table equals the number of ways to arrange 7 people in a row.

\item How many arrangements are there of the letters of TATTY?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Glue blocks, count gaps, divide by repeats. Name the technique before computing.}

\begin{enumerate}

\item Six people stand in a row; Priya and Quinn stand together. How many arrangements?

\item Six people stand in a row; Priya and Quinn are \emph{not} together. How many arrangements?

\item How many arrangements are there of the letters of MISSISSIPPI?

\item Five boys and three girls stand in a row so that no two girls are adjacent. How many arrangements?

\item Four men and four women stand in a row, alternating gender. How many arrangements?

\item Six people sit around a round table; two particular people insist on sitting together. How many seatings? (Rotations count as the same.)

\item How many six-digit strings use each of the digits $1,1,2,2,3,3$?

\item How many arrangements are there of the letters of LEVEL?

\item Seven different books are shelved in a row. Three of them form a numbered trilogy which must appear in volume order (volume 1 before 2 before 3), though not necessarily adjacently. How many shelvings?

\item The number of ways to seat $n$ people around a round table is 720. Find $n$.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Combine blocks, gaps and circles. Decide what is glued, what is free.}

\begin{enumerate}

\item Eight people stand in a row with three particular friends in consecutive positions. How many arrangements?

\item Same eight people, the three friends consecutive \emph{and} in height order (tallest first). How many arrangements now?

\item How many arrangements are there of the letters of DIVIDED?

\item Four men and four women sit around a round table, alternating gender. How many seatings? (Rotations count as the same.)

\item How many arrangements of the letters of CIRCLE have the two C's apart?

\item How many five-digit numbers use each of $1,2,3,4,5$ exactly once and are even?

\item Three maths books, two physics books and two chemistry books (all different) are shelved with books of the same subject together. How many shelvings?

\item Eight people sit around a round table; two particular people refuse to sit next to each other. How many seatings? (Rotations count as the same.)

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(SMC difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Insight over computation. If the last ones resist, sit with them --- learn the tool, then return.}

\begin{enumerate}

\item How many seven-digit numbers use each of $1,2,\ldots,7$ exactly once with the odd digits appearing in increasing order from left to right (not necessarily adjacent)?

\item Ten children stand in a circle. Mia and Noah must be adjacent, while Omar and Priya must \emph{not} be adjacent. How many circles are possible? (Rotations count as the same.)

\item How many arrangements of AABBCC have no two identical letters adjacent?

\item Eight people, including three sisters, stand in a row so that \emph{exactly} two of the sisters are adjacent. How many arrangements?

\item Eight knights sit at a round table. Three of them are feuding, and no two of the three may sit next to each other. How many seatings are possible? (Rotations count as the same.)

\end{enumerate}

\SpeedClosing{``Glue what must be together, seat what must be apart last ---
the gaps will tell you how much room rebellion really has.''}

\end{document}
