\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Combinatorics}{5}

\begin{document}

\SpeedTitleBlock{Daily Combinatorics Drill \#5}
\noindent\textit{New toolkit today: Stars \& Bars $\cdot$ Substitution Shifts $\cdot$ Inclusion--Exclusion $\cdot$ Counting with Caps.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Distributions and overlaps on sight. One formula each.}

\begin{enumerate}

\item How many pairs of non-negative integers $(x,y)$ satisfy $x+y=10$?

\item In how many ways can 5 identical sweets be shared among 3 children (a child may get none)?

\item Given $|A|=15$, $|B|=12$ and $|A\cap B|=5$, find $|A\cup B|$.

\item How many integers from 1 to 100 are divisible by 3?

\item How many integers from 1 to 100 are divisible by 3 or by 5?

\item How many triples of non-negative integers $(x,y,z)$ satisfy $x+y+z=4$?

\item How many triples of \emph{positive} integers $(x,y,z)$ satisfy $x+y+z=6$?

\item In a class of 30, twenty study French, fifteen study German and eight study both. How many study neither?

\item True or false: the number of ways to write 7 as an ordered sum of two positive integers is 6.

\item In how many ways can 4 identical marbles be put into 2 boxes?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Stars and bars, shifts, and two-set overlaps. Write the substitution down.}

\begin{enumerate}

\item How many triples of non-negative integers satisfy $x+y+z=12$?

\item How many triples of positive integers satisfy $x+y+z=12$?

\item Ten identical sweets are shared among 4 children so that each child gets at least one. In how many ways?

\item How many integers from 1 to 1000 are divisible by neither 2 nor 5?

\item How many triples of non-negative integers satisfy $x+y+z=10$ with $x\geq2$?

\item In how many arrangements of the digits $1,2,3,4$ is the digit 1 \emph{not} in the first position?

\item How many integers from 1 to 100 are divisible by 2 or by 3?

\item In how many ways can 8 be written as an ordered sum of three positive integers?

\item How many pairs of non-negative integers $(x,y)$ satisfy $x+y\leq5$?

\item How many 5-letter strings over the alphabet $\{$A, B$\}$ use both letters at least once?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Shift for lower bounds, include--exclude for upper bounds. State which you are using.}

\begin{enumerate}

\item How many quadruples of positive integers satisfy $x+y+z+w=15$?

\item How many triples of non-negative integers satisfy $x+y+z=15$ with every variable at most 6?

\item Of 100 students, 50 like tea, 40 like coffee, 30 like juice; 20 like tea and coffee, 15 tea and juice, 10 coffee and juice; 5 like all three. How many like none of the drinks?

\item How many integers from 1 to 1000 are divisible by 2, 3, or 5?

\item Eight identical apples are shared among 3 children so that no child gets more than 4. In how many ways?

\item How many arrangements of ABCDE have A not in the first position and B not in the second?

\item Three standard dice are rolled in order. How many outcomes have the three faces summing to 10?

\item How many positive integers less than 1000 have digit sum equal to 9?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(SMC / BMO1 difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Proof expected. Write the argument, not just the number.}

\begin{enumerate}

\item Four letters are placed into four addressed envelopes, one per envelope. Show, using inclusion--exclusion, that in exactly 9 placements no letter is in its correct envelope.

\item Prove that the number of ways to write $n$ as an \emph{ordered} sum of positive integers is $2^{n-1}$.

\item How many integers from 1 to 360 share no common factor with 360 other than 1? Justify via inclusion--exclusion.

\item Five students each join exactly one of three clubs, and every club must gain at least one member. Show that there are exactly 150 ways this can happen.

\item How many integers from $1$ to $10^6$ are neither perfect squares nor perfect cubes? Take care with the numbers that are both.

\end{enumerate}

\SpeedClosing{``Overcount boldly, then subtract with precision ---
inclusion--exclusion is bookkeeping for the brave.''}

\end{document}
