\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Combinatorics}{4}

\begin{document}

\SpeedTitleBlock{Daily Combinatorics Drill \#4}
\noindent\textit{New toolkit today: Binomial Theorem $\cdot$ Pascal's Rule $\cdot$ Coefficient Extraction $\cdot$ Substituting Values into Expansions.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Coefficients on sight. One binomial fact per question.}

\begin{enumerate}

\item Find the coefficient of $x^2$ in $(1+x)^5$.

\item Find the coefficient of $x$ in $(x+2)^3$.

\item Write down row 5 of Pascal's triangle.

\item Find the coefficient of $x^3y^2$ in $(x+y)^5$.

\item What is the sum of all the coefficients in the expansion of $(1+x)^7$?

\item Find the constant term of $\left(x+\dfrac{1}{x}\right)^{4}$.

\item Evaluate $\dbinom{9}{4}+\dbinom{9}{5}$ as a single binomial coefficient, and state its value.

\item Evaluate: $\dbinom{6}{0}-\dbinom{6}{1}+\dbinom{6}{2}-\cdots+\dbinom{6}{6}$.

\item True or false: the coefficient of $x^2$ in $(1-x)^{10}$ is negative.

\item Write down the largest coefficient in the expansion of $(1+x)^4$.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{General terms and clean extractions. Write $\binom{n}{k}a^{n-k}b^k$ and read off $k$.}

\begin{enumerate}

\item Find the coefficient of $x^3$ in $(2+3x)^5$.

\item Find the coefficient of $x^5$ in $(1+x)^4(1+x)^3$.

\item Find the term independent of $x$ in $\left(2x+\dfrac{1}{x^2}\right)^{6}$.

\item The coefficient of $x^2$ in $(1+x)^n$ is 66. Find $n$.

\item Evaluate: $\dbinom{10}{0}+\dbinom{10}{2}+\dbinom{10}{4}+\cdots+\dbinom{10}{10}$.

\item Find the coefficient of $x^2$ in $(1+2x)^4(1-x)^2$.

\item Use the binomial theorem to evaluate $11^3$ and $9^3$ without long multiplication.

\item The coefficient of $x^3$ in $(1+kx)^6$ is 160. Find $k$.

\item Find the value of $n$ for which $\dbinom{n}{2}=\dbinom{n}{3}$.

\item Find the coefficient of $a^2b^2c$ in $(a+b+c)^5$. \textit{\small(cf.\ TMUA 2020 P1 Q13)}

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Choose what to substitute --- $x=1$, $x=-1$, or a cunning value --- before expanding anything.}

\begin{enumerate}

\item Find the coefficient of $x^6$ in $\left(x^2+\dfrac{1}{x}\right)^{6}$.

\item Which coefficient in the expansion of $(1+x)^8$ is the greatest, and what is its value?

\item Use the first three terms of a binomial expansion to estimate $1.01^{10}$ to 4 decimal places.

\item The coefficients of $x^4$ and $x^5$ in $(1+x)^n$ are equal. Find $n$.

\item Find the coefficient of $x^4$ in $(1+x)^5(1-2x)^3$. \textit{\small(cf.\ MAT 2023 Q1F)}

\item Evaluate $\;\displaystyle\sum_{k=0}^{5}k\binom{5}{k}$.

\item Find the constant term of $\left(x-\dfrac{2}{x}\right)^{8}$.

\item Prove Pascal's rule $\dbinom{n}{r}=\dbinom{n-1}{r-1}+\dbinom{n-1}{r}$ by a committee argument, and use it to generate row 7 of Pascal's triangle from row 6.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(SMC / BMO1 difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Proof enters the drill. Write your arguments in full sentences --- BMO1 marks the writing, not the answer.}

\begin{enumerate}

\item Prove that $\dbinom{2n}{n}$ is even for every integer $n\geq1$.

\item Show that $\;\displaystyle\sum_{k=0}^{10}\binom{10}{k}^{2}=\binom{20}{10}$, by counting one scenario in two ways.

\item For how many integers $k$ with $0\leq k\leq12$ is $\dbinom{12}{k}$ odd? \textit{\small(cf.\ TMUA 2023 P1 Q6)}

\item Show that $\;\dbinom{2}{2}+\dbinom{3}{2}+\dbinom{4}{2}+\cdots+\dbinom{9}{2}=\dbinom{10}{3}$, and state its value.

\item Let $p$ be a prime and $0<k<p$. Prove that $\dbinom{p}{k}$ is divisible by $p$.

\end{enumerate}

\SpeedClosing{``An identity is a story told twice ---
once for each side of the equals sign.''}

\end{document}
