\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Combinatorics}{2}

\begin{document}

\SpeedTitleBlock{Daily Combinatorics Drill \#2}
\noindent\textit{New toolkit today: Combinations $\binom{n}{r}$ $\cdot$ Choosing vs Arranging $\cdot$ Complementary Counting $\cdot$ Committee Constructions.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Evaluate or count instantly. Choosing, not arranging. Pure recognition.}

\begin{enumerate}

\item Evaluate: $\;\dbinom{7}{2}$.

\item Evaluate: $\;\dbinom{8}{3}$.

\item Evaluate: $\;\dbinom{9}{7}$.

\item How many committees of 3 can be chosen from 6 people?

\item Solve for $n$: $\;\dbinom{n}{2}=45$.

\item How many subsets of a 5-element set have at least 4 elements?

\item Evaluate: $\;\dbinom{6}{0}+\dbinom{6}{1}+\dbinom{6}{2}+\cdots+\dbinom{6}{6}$.

\item How many ways can 2 letters be chosen from $\{$A, B, C, D, E$\}$?

\item True or false: $\;\dbinom{100}{98}=4950$.

\item A lottery ticket selects 4 numbers from $1$--$10$ (order irrelevant). How many different tickets are there?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Set up the binomial coefficient, don't list. Watch for choose-then-choose products.}

\begin{enumerate}

\item Solve for $n$: $\;\dbinom{n}{3}=35$.

\item From 8 men and 6 women, how many committees of 2 men and 2 women can be formed?

\item Twelve people at a party all shake hands with each other exactly once. How many handshakes take place?

\item Ten points lie in the plane, no three collinear. How many line segments join pairs of them, and how many triangles have all three vertices among them?

\item A team of 5 is chosen from a squad of 11, and then one of the 5 is named captain. How many ways? Check your answer by choosing the captain first.

\item How many binary strings of length 8 contain exactly three 1s?

\item How many 3-letter strings (26 letters, repeats allowed) contain at least one vowel (A, E, I, O, U)?

\item Solve for $n$: $\;\dbinom{n}{2}+\dbinom{n}{1}=21$.

\item A committee of 4 is chosen from 9 people. How many committees include Priya? How many exclude her?

\item A robot walks from $(0,0)$ to $(4,3)$ taking unit steps only right or up. How many routes are there?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Case splits and complements. Decide "at least / at most / exactly" before computing.}

\begin{enumerate}

\item A committee of 5 is chosen from 7 women and 6 men and must contain at least 4 women. How many committees are possible?

\item How many four-digit numbers have at least one repeated digit?

\item How many diagonals does a regular 12-gon have?

\item Eight points lie on a circle. How many triangles have all vertices among them? How many of those triangles use the point $A$?

\item From 6 boys and 6 girls, a team of 6 with equal numbers of boys and girls is chosen. How many teams?

\item How many 5-card hands from a standard 52-card deck contain all four aces?

\item A convex $n$-gon has exactly 27 diagonals. Find $n$.

\item How many subsets of $\{1,2,\ldots,10\}$ contain at least one of the elements 1 and 2?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(SMC difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Insight over computation. If the last ones resist, sit with them --- learn the tool, then return.}

\begin{enumerate}

\item How many rectangles (of any size) can be drawn on a $4\times4$ grid of unit squares with sides along the gridlines? \textit{\small(extends Sheet 1, D2)}

\item How many subsets of $\{1,2,\ldots,8\}$ contain no two consecutive integers? (Count the empty set.)

\item A club has 7 juniors and 5 seniors. How many committees of 4 have strictly more juniors than seniors?

\item Six people each shake hands with exactly two of the others. In how many ways can this happen?

\item Ten points lie on a circle, and every pair is joined by a chord. If no three chords meet at a single interior point, how many interior intersection points are there?

\end{enumerate}

\SpeedClosing{``When a count splits into cases, count the complement first ---
the opposite of `at least one' is always simpler.''}

\end{document}
