\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Logic}{7}

\begin{document}

\SpeedTitleBlock{Daily Logic Drill \#7}{David Akinyele-Aje}
\noindent\textit{New toolkit today: Proof by Contradiction $\cdot$ Combining Techniques $\cdot$ Synoptic Mastery.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{The mechanics of proof by contradiction. Pure recognition.}

\begin{enumerate}

\item To prove statement $P$ by contradiction, you should assume \_\_\_ and derive a contradiction.

\item True or false: ``Proof by contradiction and proof by contrapositive are the same technique.''

\item What contradiction is reached in the classical proof that $\sqrt2$ is irrational?

\item True or false: ``If assuming `not $P$' leads to a statement that's always false, $P$ must be true.''

\item Give the first line you would write to prove ``there are infinitely many primes'' by contradiction.

\item True or false: ``Deriving `$1=0$' partway through an argument is a valid contradiction to conclude the assumption was false.''

\item True or false: ``In a proof by contradiction of `if $P$ then $Q$', you assume $P$ is true and $Q$ is false simultaneously.''

\item True or false: ``Every valid proof by contrapositive can be rewritten as a proof by contradiction.''

\item To disprove a universal (``for all'') claim as fast as possible, should you generally reach for a counterexample or for proof by contradiction?

\item True or false: ``If you can find a single counterexample to a claim, you never need a separate proof by contradiction to disprove that same claim.''

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Write short, complete proofs by contradiction.}

\begin{enumerate}

\item Prove by contradiction: ``There is no smallest positive rational number.''

\item Prove by contradiction: ``If $n^2$ is even, then $n$ is even.''

\item Prove by contradiction: ``There is no largest even integer.''

\item Prove by contradiction: ``If $a,b$ are integers and $a+b$ is odd, then $a$ and $b$ cannot both be even.''

\item Prove by contradiction: ``If $p$ is prime and $p>3$, then $p$ is not divisible by $6$.''

\item Prove by contradiction: ``There is no integer $n$ satisfying both $n>5$ and $n<3$.''

\item Prove by contradiction: ``If $x$ is a positive real number, then $x+\dfrac1x \geq 2$.''

\item Prove by contradiction: ``If $n$ is a positive integer, $n$ and $n+1$ share no common factor greater than $1$.''

\item Prove by contradiction: ``$\log_2 3$ is irrational.''

\item Prove by contradiction: ``If $a,b$ are positive integers with $a<b$, then $a^2<b^2$.''

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Longer contradiction proofs, combining tools from across the whole pillar.}

\begin{enumerate}

\item Prove by contradiction: ``$\sqrt2+\sqrt3$ is irrational.''

\item Prove by contradiction: ``There is no integer solution to $x^2+y^2=4z+3$.''

\item You are given the identity $2^n-1=(2^a-1)(2^{n-a}+2^{n-2a}+\cdots+1)$ whenever $a$ divides $n$. Using proof by contradiction, show that if $2^n-1$ is prime then $n$ is prime.

\item Prove by contradiction: ``If $n$ is a positive integer that is not a perfect square, then $\sqrt n$ is irrational.''

\item Using proof by contradiction and the fact that $2$ is the only even prime (Day 1, B9), prove there do not exist two distinct primes $p,q$, both greater than $2$, with $p+q=5$.

\item Using contradiction and the fact that $n^2-n$ is always even (Day 1, D5 / Day 6, C5), prove there is no positive integer $n$ for which $n^2-n=7$.

\item Prove by contradiction: ``The equation $x^2+y^2=3$ has no solution in positive integers.''

\item Using contradiction, prove that if $a,b,c$ are positive integers with $a^2+b^2=c^2$ and $c$ is even, then $a$ and $b$ cannot both be odd.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(TMUA / SMC difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Insight over computation. Each question has a short, elegant solution --- find it.}

\begin{enumerate}

\item Consider: I. ``Proof by contradiction and proof by contrapositive always produce identical proofs.'' II. ``A single counterexample is a special case of proof by contradiction.'' III. ``If a statement's negation leads to `$1=0$', the statement is proven true.'' Which are true?
\begin{itemize}
\item[A)] I only
\item[B)] II only
\item[C)] III only
\item[D)] I and III only
\item[E)] I, II and III
\end{itemize}

\item Prove, using proof by contradiction, that there is no positive integer $n$ for which both $n$ and $n+1$ are perfect squares.

\item In a room of at least $2$ people, every person is either a knight (always tells the truth) or a knave (always lies). Every person in the room says: ``Everyone else in this room is a knave.'' Using proof by contradiction, determine exactly how many knights are in the room.

\item Using proof by contradiction, prove that the equation $n^2=2^k$ has no solution in positive integers $n,k$ with $k$ odd.

\item Give a second, different proof that $\sqrt2$ is irrational (Day 5, D2 gave one proof already), this time using the same $2$-adic valuation technique as D4.

\end{enumerate}

\SpeedClosing{``Speed comes from recognition, not from rushing.
Every shortcut is a pattern you have internalised.''}

\end{document}
