\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Logic}{6}

\begin{document}

\SpeedTitleBlock{Daily Logic Drill \#6}{David Akinyele-Aje}
\noindent\textit{New toolkit today: Evaluating Compound Claims $\cdot$ AND/OR Combinations $\cdot$ The ``Which Statements Are True'' Format $\cdot$ Quantifier Mixing.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Evaluate each statement independently before combining. Pure recognition.}

\begin{enumerate}

\item I. ``$2$ is prime.'' II. ``$2$ is odd.'' Which are true?

\item I. ``$4$ is a perfect square.'' II. ``$4$ is prime.'' Which are true?

\item I. ``$9$ is odd.'' II. ``$9$ is prime.'' Which are true?

\item I. ``Every square is a rectangle.'' II. ``Every rectangle is a square.'' Which are true?

\item I. ``$6$ is a multiple of $2$.'' II. ``$6$ is a multiple of $3$.'' Which are true?

\item I. ``$5$ is even.'' II. ``$5$ is prime.'' Which are true?

\item I. ``$0$ is positive.'' II. ``$0$ is negative.'' Which are true?

\item True or false: ``If both statements I and II in a pair are false, the correct choice is `neither', not `I and II'.''

\item I. ``$n=2$ is a counterexample to `all primes are odd'.'' II. ``$n=9$ is a counterexample to `all odd numbers are prime'.'' Which are true?

\item I. ``$1$ is a prime number.'' II. ``$1$ is a perfect square.'' Which are true?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Three statements at once --- evaluate each on its own merits before combining.}

\begin{enumerate}

\item I. ``$n=4$ is a multiple of $4$.'' II. ``$n=4$ is a multiple of $8$.'' III. ``$n=4$ is a multiple of $2$.'' Which are true?

\item I. ``$n$ even $\Rightarrow n^2$ even'' is true. II. Its converse is also true. III. Together, I and II mean ``$n$ even $\iff n^2$ even.'' Which are true?

\item I. ``$\sqrt2$ is irrational.'' II. ``$\sqrt2+\sqrt2$ is irrational.'' III. ``$\sqrt2-\sqrt2$ is irrational.'' Which are true?

\item I. ``Every multiple of $9$ is a multiple of $3$.'' II. ``Every multiple of $3$ is a multiple of $9$.'' III. ```Multiple of $3$' is necessary for `multiple of $9$'.'' Which are true?

\item I. ``$2^{11}-1$ is prime.'' II. ``$11$ is prime.'' III. ``If $n$ is prime, $2^n-1$ is prime.'' Which are true?

\item I. ``$4!+1$ is prime.'' II. ``$5!+1$ is prime.'' III. ``$n!+1$ is prime for all $n$.'' Which are true?

\item I. ``$41$ is prime.'' II. ``$n=41$ is a counterexample to `$n^2+n+41$ is always prime'.'' III. ``$n=40$ is the smallest counterexample to `$n^2+n+41$ is always prime'.'' Which are true?

\item I. ``The negation of `for all $x$, $P(x)$' is `for all $x$, not $P(x)$'.'' II. ``The negation of `for all $x$, $P(x)$' is `there exists $x$, not $P(x)$'.'' III. ``The negation of `there exists $x$, $P(x)$' is `for all $x$, not $P(x)$'.'' Which are true?

\item I. ``The converse of a true statement is always true.'' II. ``The contrapositive of a true statement is always true.'' III. ``The negation of a true statement is always true.'' Which are true?

\item I. ``Affirming the consequent is a valid rule.'' II. ``Denying the antecedent is a valid rule.'' III. ``Modus ponens is a valid rule.'' Which are true?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Harder combinations, drawing on facts from across the whole pillar so far.}

\begin{enumerate}

\item For a positive integer $n$: I. ``$n$ divisible by $4$'' is necessary for ``$n$ divisible by $12$.'' II. ``$n$ divisible by $4$'' is sufficient for ``$n$ divisible by $12$.'' III. ``$n$ divisible by $4$'' is necessary and sufficient for ``$n$ divisible by $8$.'' Which of I, II, III are true?
\begin{itemize}
\item[A)] I only
\item[B)] II only
\item[C)] III only
\item[D)] I and II only
\item[E)] I and III only
\item[F)] II and III only
\item[G)] I, II and III
\item[H)] None of them
\end{itemize}

\item I. ``There exists a real $x$ with $x^2<0$.'' II. ``For all reals $x$, $x^2\geq0$.'' III. ``The negation of II is I.'' Which are true?

\item Recall the flawed argument ``$n^2$ multiple of $4$, therefore $n$ multiple of $4$'' (Day 5, B1). I. ``This argument commits the fallacy of affirming the consequent.'' II. ``$n=6$ is a valid counterexample to the underlying claim.'' III. ``The converse of the underlying claim (`$n$ mult.\ of $4 \Rightarrow n^2$ mult.\ of $4$') is what's actually true.'' Which are true?

\item I. ``$n=1$ is a counterexample to `every positive integer has a prime factor'.'' II. ``$n=1$ is a counterexample to `every integer greater than $1$ has a prime factor'.'' III. ``$1$ is neither prime nor composite.'' Which are true?

\item I. ``The statement `if $n^2-n$ is even, then $n$ is even' is false (Day 1, D5).'' II. ``$n^2-n$ is even for every integer $n$.'' III. ``$n(n-1)$ is even for every integer $n$.'' Which are true?

\item I. ``Every quadratic with integer coefficients that has an integer root has two integer roots.'' II. ``$2x^2-3x+1$ is a counterexample to I.'' III. ``Monic quadratics (leading coefficient $1$) are never counterexamples to I.'' Which are true?

\item I. ``$3^5+2$ is composite (Day 4, C6).'' II. ``$3^5+2=245$.'' III. ``$245=5\times7^2$.'' Which are true?

\item For real $x$: I. ``$x>0$'' is sufficient for ``$x^2>0$.'' II. ``$x>0$'' is necessary for ``$x^2>0$.'' III. ``$x\neq0$'' is necessary and sufficient for ``$x^2>0$.''
\begin{itemize}
\item[A)] I only
\item[B)] I and II only
\item[C)] I and III only
\item[D)] II and III only
\item[E)] I, II and III
\end{itemize}

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(TMUA / SMC difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Insight over computation. Each question has a short, elegant solution --- find it.}

\begin{enumerate}

\item For a positive integer $n$: I. ``$n$ divisible by $6$'' is necessary and sufficient for ``$n$ divisible by $2$ and by $3$.'' II. ``$n^2$ divisible by $9$'' is necessary and sufficient for ``$n$ divisible by $3$.'' III. ``$n^2$ divisible by $4$'' is necessary and sufficient for ``$n$ divisible by $4$.'' Which of I, II, III are true?
\begin{itemize}
\item[A)] I only
\item[B)] II only
\item[C)] III only
\item[D)] I and II only
\item[E)] I and III only
\item[F)] II and III only
\item[G)] I, II and III
\item[H)] None of them
\end{itemize}

\item Three people, $X$, $Y$, $Z$, each make one statement. $X$ says: ``$Y$ is lying.'' $Y$ says: ``$Z$ is lying.'' $Z$ says: ``$X$ and $Y$ are both lying.'' Given that exactly one of them is telling the truth, determine who.

\item Recall the flawed conclusion from Day 5, D2 (a student proves $\sqrt2$ irrational by contradiction, then writes ``therefore $\sqrt2$ is rational''). I. ``The algebraic derivation, up to the final sentence, is entirely correct.'' II. ``The final English conclusion correctly follows from the derived contradiction.'' III. ``A correctly-stated conclusion would be `$\sqrt2$ is irrational'.'' Which are true?

\item Let $S$ be the statement: ``If $n$ is a multiple of $12$, then $n$ is a multiple of $3$ and a multiple of $4$.'' I. $S$ is true. II. $S$'s converse is true. III. ``Multiple of $12$'' is necessary and sufficient for ``multiple of $6$.'' Which are true?

\item Statement $T$: ``If $2^n-1$ is prime, then $n$ is prime'' (Day 4, D5). I. $T$'s contrapositive is ``if $n$ is not prime, then $2^n-1$ is not prime.'' II. $T$'s converse is false, witnessed by $n=11$. III. ``$2^n-1$ is prime'' is necessary and sufficient for ``$n$ is prime.'' Which are true?

\end{enumerate}

\SpeedClosing{``Speed comes from recognition, not from rushing.
Every shortcut is a pattern you have internalised.''}

\end{document}
