\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Logic}{1}

\begin{document}

\SpeedTitleBlock{Daily Logic Drill \#1}{David Akinyele-Aje}
\noindent\textit{New toolkit today: Truth Values \& Statements $\cdot$ If--Then Structure $\cdot$ Converse vs Original $\cdot$ Basic Direct Proof.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Hypothesis, conclusion, converse. Pure recognition, no working.}

\begin{enumerate}

\item Identify the hypothesis and the conclusion of: ``If $n$ is divisible by $10$, then $n$ is divisible by $5$.''

\item State the converse of: ``If $x = 3$, then $x^2 = 9$.''

\item True or false: ``If a shape is a square, then it is a rectangle.''

\item True or false: ``If a shape is a rectangle, then it is a square.''

\item State the converse of: ``If $n$ is even, then $n^2$ is even.''

\item True or false: ``$n^2 > 0$ for every integer $n$.''

\item In ``If $x>3$ then $x^2 > 9$'', which part is the hypothesis and which is the conclusion?

\item True or false: ``If $a=b$, then $a^2 = b^2$.''

\item State the converse of: ``If $n$ is a multiple of $6$, then $n$ is a multiple of $2$.''

\item True or false: ``The converse of a true statement is always true.''

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Build the converse of a compound statement, or write a short direct proof.}

\begin{enumerate}

\item Write the converse of: ``If $n$ is divisible by $4$ and by $3$, then $n$ is divisible by $12$.'' State whether the original statement and its converse are both true.

\item Prove directly: ``If $n$ is an even integer, then $n^2$ is even.''

\item State the converse of: ``If $x>0$ and $y>0$, then $xy>0$.'' Is the converse true? Justify your answer.

\item True or false, with a one-line reason: ``If a number's digits sum to a multiple of $3$, then the number is a multiple of $3$.''

\item Prove directly: ``If $n$ is an odd integer, then $n^2$ is odd.''

\item Consider: ``If $x^2 = 9$ then $x = 3$.'' Is this true? If false, give the value of $x$ that breaks it.

\item State the converse of Pythagoras' theorem (``if a triangle has a right angle, then $a^2+b^2=c^2$ for its side lengths, with $c$ the longest''). Is this converse also a true geometric fact?

\item Prove directly: ``If $a$ and $b$ are both even integers, then $a+b$ is even.''

\item True or false, with a one-line reason: ``If $p$ is a prime number and $p \neq 2$, then $p$ is odd.''

\item State the converse of: ``If $x=y$, then $x^3 = y^3$.'' Is the converse true?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Converses that hide a false step, proofs that need one extra insight, and your first multiple-choice question.}

\begin{enumerate}

\item Prove directly that the square of any multiple of $3$ is itself a multiple of $9$.

\item Prove directly: ``If $n$ is an odd integer, then $n^2 - 1$ is divisible by $8$.''

\item Which of the following is the correct converse of: ``If $x$ is rational and $y$ is irrational, then $x+y$ is irrational''?
\begin{itemize}
\item[A)] If $x+y$ is irrational, then $x$ is rational and $y$ is irrational.
\item[B)] If $x$ is irrational and $y$ is rational, then $x+y$ is irrational.
\item[C)] If $x+y$ is rational, then $x$ is irrational or $y$ is rational.
\item[D)] If $x$ is rational and $y$ is rational, then $x+y$ is irrational.
\end{itemize}

\item True or false, with justification: ``If a statement is true, its converse must also be false.''

\item Prove directly: ``If $n$ is divisible by $6$, then $n^3$ is divisible by $216$.''

\item State the converse of: ``For positive reals $x,y$: if $x \geq 2$ and $y \geq 2$, then $xy \geq x+y$.'' Determine whether the converse is true, giving a counterexample if it is false.

\item State the converse of: ``If $x$ and $y$ are both perfect squares, then $xy$ is a perfect square.'' Is the converse true? Give a counterexample if it is false.

\item True or false, with justification: ``If $n^2$ is a multiple of $4$, then $n$ is a multiple of $4$.''

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(TMUA / SMC difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Insight over computation. Each question has a short, elegant solution --- find it.}

\begin{enumerate}

\item Let $n$ be a positive integer. Consider the statements:
\begin{itemize}
\item[I.] If $n$ is divisible by $4$, then $n$ is divisible by $8$.
\item[II.] If $n$ is divisible by $8$, then $n$ is divisible by $4$.
\item[III.] The converse of statement I is statement II.
\end{itemize}
Which of the statements I, II, III are true?
\begin{itemize}
\item[A)] I only
\item[B)] II only
\item[C)] III only
\item[D)] I and II only
\item[E)] I and III only
\item[F)] II and III only
\item[G)] I, II and III
\item[H)] None of them
\end{itemize}

\item Prove that if $n$ and $n+2$ are both prime numbers greater than $3$, then $n+1$ is divisible by $6$.

\item Amy claims: ``If a positive integer is divisible by every one of $2,3,4,5,6$, then it is divisible by $60$.'' Determine whether Amy's claim is true, and justify your answer using the smallest positive integer for which it can be checked.

\item Give an example of a true statement $S$ about divisibility of integers whose converse $S'$ is also true. State both $S$ and $S'$.

\item Determine, with proof, whether the following statement is true or false: ``If $n^2 - n$ is even, then $n$ is even.''

\end{enumerate}

\SpeedClosing{``Speed comes from recognition, not from rushing.
Every shortcut is a pattern you have internalised.''}

\end{document}
