\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Logic}{2}

\begin{document}

\SpeedTitleBlock{Daily Logic Drill \#2}{David Akinyele-Aje}
\noindent\textit{New toolkit today: Necessary vs Sufficient Conditions $\cdot$ ``Iff'' $\cdot$ Chains of Implication $\cdot$ Weakest/Strongest Condition.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Translate familiar if--then facts into necessary/sufficient language. Pure recognition.}

\begin{enumerate}

\item Is ``$n$ is divisible by $4$'' a sufficient condition for ``$n$ is divisible by $2$''?

\item Is ``$n$ is divisible by $4$'' a necessary condition for ``$n$ is divisible by $2$''?

\item Is ``$x=3$'' a sufficient condition for ``$x^2=9$''?

\item Is ``$x=3$'' a necessary condition for ``$x^2=9$''?

\item True or false: ``being a rectangle'' is a necessary condition for ``being a square''.

\item True or false: ``being a rectangle'' is a sufficient condition for ``being a square''.

\item Is ``$n=6$'' a sufficient condition for ``$n$ is a multiple of $3$''?

\item Is ``$n=6$'' a necessary condition for ``$n$ is a multiple of $3$''?

\item True or false: ``$n$ is even'' is a necessary condition for ``$n$ is divisible by $4$''.

\item True or false: ``$n$ is even'' is a sufficient condition for ``$n$ is divisible by $4$''.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Classify a condition as necessary, sufficient, both, or neither; chain implications together.}

\begin{enumerate}

\item Classify ``$x>5$'' as a condition for ``$x>0$'': necessary, sufficient, both, or neither?

\item Classify ``$n$ is divisible by $6$'' as a condition for ``$n$ is divisible by $2$ and by $3$''.

\item Classify ``$ab=0$'' as a condition for ``$a=0$'' (where $a,b$ are real numbers).

\item Classify ``$n^2$ is even'' as a condition for ``$n$ is even''.

\item Classify ``$x$ is an integer'' as a condition for ``$x$ is a rational number''.

\item Classify ``a triangle is equilateral'' as a condition for ``a triangle is isosceles''.

\item Given that ``$P \Rightarrow Q$'' and ``$Q \Rightarrow R$'' are both true, is $P$ a sufficient condition for $R$? Justify briefly.

\item Given that ``$n$ divisible by $9 \Rightarrow n$ divisible by $3$'' and ``$n$ divisible by $3 \Rightarrow$ digit sum divisible by $3$'', is ``$n$ divisible by $9$'' sufficient for ``digit sum divisible by $3$''?

\item Each of (a) $n$ divisible by $8$, (b) $n$ divisible by $4$, (c) $n$ divisible by $2$ is sufficient for ``$n$ is even''. Which is the \emph{weakest} (least restrictive) of the three?

\item Each of (a) $n$ divisible by $2$, (b) $n$ divisible by $4$, (c) $n$ divisible by $8$ is necessary for ``$n$ is divisible by $24$''. Which is the \emph{strongest} (most restrictive) of the three?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Full necessary-and-sufficient proofs, implication chains, and your first four-way classification.}

\begin{enumerate}

\item Which of the following best describes ``$n$ is a multiple of $6$'' as a condition for ``$n$ is a multiple of $2$ and a multiple of $3$''?
\begin{itemize}
\item[A)] Necessary but not sufficient.
\item[B)] Sufficient but not necessary.
\item[C)] Necessary and sufficient.
\item[D)] Neither necessary nor sufficient.
\end{itemize}

\item Prove that ``$n$ is divisible by $4$'' is necessary and sufficient for ``$n$ is even and $n/2$ is even''.

\item Classify ``$x^2<4$'' as a condition for ``$x<2$'' (real $x$). Justify both directions.

\item Classify ``$n$ is prime'' as a condition for ``$n$ has exactly two positive divisors''. Justify your answer.

\item Given that $P$ is sufficient for $Q$, and $Q$ is sufficient for $R$, prove that $P$ is sufficient for $R$.

\item Is ``$n$ is a multiple of $4$'' necessary for ``$n$ is a multiple of $12$''? Is it sufficient? Justify both answers.

\item Find a condition $C$ on a positive integer $n$ that is necessary but not sufficient for ``$n$ is a multiple of $100$''. Justify your choice.

\item Given that $x$ is necessary and sufficient for $y$, and $y$ is necessary and sufficient for $z$, prove that $x$ is necessary and sufficient for $z$.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(TMUA / SMC difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Insight over computation. Each question has a short, elegant solution --- find it.}

\begin{enumerate}

\item For real $x$, let $P$ be ``$x>0$'' and $Q$ be ``$x^2>0$''. Is $P$ necessary and sufficient for $Q$, necessary but not sufficient, sufficient but not necessary, or neither?
\begin{itemize}
\item[A)] Necessary and sufficient.
\item[B)] Necessary but not sufficient.
\item[C)] Sufficient but not necessary.
\item[D)] Neither necessary nor sufficient.
\end{itemize}

\item Classify ``$x$ is rational'' as a condition for ``$x^2$ is rational'' (real $x$). Prove both parts of your classification.

\item A year $Y$ is a leap year exactly when $Y$ is divisible by $4$, except that century years (multiples of $100$) must be divisible by $400$. Is ``$Y$ is divisible by $4$'' necessary and sufficient, necessary but not sufficient, sufficient but not necessary, or neither, for ``$Y$ is a leap year''? Justify with a specific year.

\item Statements $A$, $B$, $C$ satisfy $A \Rightarrow B$, $B \Rightarrow C$, and $C \Rightarrow A$. Prove that $A$, $B$, $C$ are all logically equivalent (i.e.\ each pair is an ``iff'').

\item Statement $S$ is: ``If $n$ is a multiple of $9$, then $n$ is a multiple of $3$.'' Is ``$n$ is a multiple of $9$'' necessary and sufficient for ``$n$ is a multiple of $3$''? If not, precisely which of necessary/sufficient does it satisfy, and what does $S$'s converse (from Day 1's toolkit) have to do with your answer?

\end{enumerate}

\SpeedClosing{``Speed comes from recognition, not from rushing.
Every shortcut is a pattern you have internalised.''}

\end{document}
