\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Number Theory}{3}

\begin{document}

\SpeedTitleBlock{Daily Number Theory Drill \#3}{David Akinyele-Aje}
\noindent\textit{New toolkit today: Cyclic Patterns $\cdot$ Telescoping Exponents $\cdot$ Last Digit Sequences.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Remainders, parity, basic prime factors, and pure recognition.}

\begin{enumerate}

\item Find the remainder when $10^{10} + 10^5 + 1$ is divided by $9$.

\item What is the unit digit of $3^{2026}$?

\item Find the largest prime factor of $1001$.

\item Find the remainder when $7 \times 8 \times 9$ is divided by $5$.

\item If $n$ is an integer such that $2n \equiv 3 \pmod 5$, what is the smallest positive value of $n$?

\item Evaluate $\gcd(144, 84)$.

\item Find the number of positive divisors of $36$.

\item Find the remainder when $2^{10}$ is divided by $11$.

\item Express $120$ as a product of prime factors.

\item Find the smallest positive integer $k$ such that $12k$ is a perfect square.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Divisibility rules, modulo properties, algebraic factoring over integers.}

\begin{enumerate}

\item How many of the numbers $3, 4, 5, 6, 7$ are factors of the sum $3^{2026} + 3^{2025} + 3^{2024}$? \textit{\small(after SMC 2023 Q10)}

\item Find the highest power of $2$ that divides $100!$.

\item If $a$ and $b$ are non-zero digits, find the maximum possible integer value of $\dfrac{10a+b}{a+b}$.

\item Find the number of integer pairs $(x,y)$ such that $xy = 24$.

\item What is the smallest positive integer greater than $2$ that leaves a remainder of $2$ when divided by $3, 4$, and $5$?

\item Find the remainder when $11^{11} + 12^{12}$ is divided by $10$.

\item Find the value of $p$ if $p$ and $p^2+2$ are both prime numbers.

\item Given that $n$ is an integer, for how many values of $n$ is $\dfrac{2n+15}{n+2}$ an integer?

\item Find the sum of all positive integers $n$ for which $n^2-1$ is a prime number.

\item Determine the two last digits of $5^{2026} + 6^{2026}$.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Link Number Theory to Algebra and Combinatorics. Look for patterns.}

\begin{enumerate}

\item Find all pairs of positive integers $(x, y)$ such that $\dfrac{1}{x} + \dfrac{1}{y} = \dfrac{1}{3}$.

\item Solve the equation $x^2 - y^2 = 2025$ for positive integers $(x,y)$ such that $y$ is as small as possible.

\item Find all pairs of integers $(x,y)$ that satisfy $xy + x + y = 14$.

\item Find all integers $x$ such that $x^4 + x^2 + 1$ is a prime number.

\item How many ways can we choose the digits $a$ and $b$ such that the 5-digit number $3a4b2$ is divisible by $36$?

\item Find the number of 3-element subsets of $\{1, 2, \dots, 10\}$ such that the sum of the elements is divisible by $3$. \textit{\small(after MAT combinatorics)}

\item Find the sum of all integer values of $x$ for which $\dfrac{x^2+7x+2}{x+1}$ is an integer.

\item Find the last two digits of $2019^{2025}$. \textit{\small(after SMC binomial last digits)}

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(TMUA / SMC / BMO1 difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Insight over computation. Each question has a short, elegant solution --- find it.}

\begin{enumerate}

\item Find the number of pairs of positive integers $(a,b)$ with $a \le b$ such that $\text{lcm}(a,b) = 2^3 \cdot 3^2 \cdot 5$. \textit{\small(after BMO1 combinatorics)}

\item Find all pairs of prime numbers $(p,q)$ such that $p^2 - 2q^2 = 1$. \textit{\small(after BMO1 Diophantine)}

\item Find the number of integer solutions to $x^3 + 2y^3 = 4z^3$. \textit{\small(after BMO1 infinite descent)}

\item Find the last three digits of $7^{9999}$. \textit{\small(after SMC massive exponents)}

\item Find the last three digits of $2025^{2026}$. \textit{\small(after SMC cyclic endings)}

\end{enumerate}

\SpeedClosing{``Speed comes from recognition, not from rushing.
Every shortcut is a pattern you have internalised.''}

\end{document}
