\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Number Theory}{5}

\begin{document}

\SpeedTitleBlock{Daily Number Theory Drill \#5}{Antigravity}
\noindent\textit{New toolkit today: Bounding Diophantine Equations $\cdot$ Inequality Logic $\cdot$ Factor Searches.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Last digits, primes, and basic properties. Pure recognition.}

\begin{enumerate}

\item Find the last digit of $3^{2025}$.

\item Factorise $1001$ completely into primes.

\item Find $\gcd(42, 105)$.

\item Find the number of positive divisors of $120$.

\item Evaluate $5^{10} \pmod 6$.

\item Find the smallest positive integer $x$ such that $2x \equiv 1 \pmod 5$.

\item Find the smallest integer $n > 1$ such that $n \equiv 1 \pmod 3$ and $n \equiv 1 \pmod 4$.

\item Write the prime factorisation of $360$.

\item Calculate $7^3 \pmod{10}$.

\item Find the largest prime factor of $9999$.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Modular arithmetic, exponentiation, divisibility rules. Look for patterns.}

\begin{enumerate}

\item The greatest power of $7$ which is a factor of $50!$ is $7^k$. What is $k$? \textit{\small(after SMC 2023 Q12)}

\item Find the smallest positive integer $n$ such that $15n$ is a perfect cube.

\item Find the remainder when $1! + 2! + 3! + \cdots + 100!$ is divided by $12$.

\item Which is larger, $3^{40}$ or $4^{30}$?

\item Find the smallest positive integer $x$ such that $3x \equiv 4 \pmod 7$.

\item Find the last two digits of $7^{2022}$.

\item How many positive factors of $1000$ are perfect squares?

\item Find all integers $x$ such that $0 \le x < 8$ and $x^2 \equiv 1 \pmod 8$.

\item Given $2^x 3^y = 12^3$, find $x+y$.

\item Find all pairs of positive integers $(x,y)$ such that $xy = x + y + 3$.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90 s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Synoptic links: use algebraic identities and combinatorial counting to solve these number theory problems.}

\begin{enumerate}

\item The expression $\frac{n^2+3n+5}{n+1}$ takes integer values for certain positive integer values of $n$. What is the sum of all such positive integer values of $n$? \textit{\small(after SMC 2023 Q19)}

\item How many of the numbers 6, 7, 8, 9, 10 are factors of the sum $2^{2024} + 2^{2023} + 2^{2022}$? \textit{\small(after SMC 2023 Q10)}

\item Find all pairs of positive integers $(x, y)$ such that $x^2 - y^2 = 17$.

\item How many ordered pairs of positive integers $(a, b)$ satisfy $\operatorname{lcm}(a, b) = 300$?

\item How many positive integers $n \le 100$ have exactly 3 positive divisors?

\item Find the sum of the digits of the largest integer $n$ such that $n^3 + 100$ is a multiple of $n+10$.

\item Find all triples of positive integers $(x,y,z)$ such that $x \le y \le z$ and $xyz = x + y + z$. \textit{\small(after SMC 2022 Q21)}

\item Find the sum of all integer values of $n$ for which $n^2 + 4n + 3$ is a perfect square. \textit{\small(after AI Prompts D5)}

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(TMUA / SMC difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Insight over computation. Each question has a short, elegant solution --- find it.}

\begin{enumerate}

\item Find all prime numbers $p$ such that $p^2 + 2$ is also a prime number.

\item Find all pairs of positive integers $(x,y)$ such that $x^2 y = x^2 + 3y + 2$. \textit{\small(after BMO1 1995 Q1)}

\item How many pairs of positive integers $(x,y)$ with $x \le y$ satisfy $\frac{1}{x} + \frac{1}{y} = \frac{1}{6}$? \textit{\small(after SMC 2021 Q24)}

\item Find the remainder when $1^{2025} + 2^{2025} + \cdots + 10^{2025}$ is divided by 11.

\item Find the sum of all integers $n$ for which $n^4 + 4$ is a prime number.

\end{enumerate}

\SpeedClosing{``Number theory is the queen of mathematics.'' --- Carl Friedrich Gauss}

\end{document}
