\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Number Theory}{1}

\begin{document}

\SpeedTitleBlock{Daily Number Theory Drill \#1}{David Akinyele-Aje}
\noindent\textit{New toolkit today: Core Divisibility $\cdot$ Parity Arguments $\cdot$ Alternating Sums.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Prime factors, parity, modular remainders. Pure recognition.}

\begin{enumerate}

\item Factorise completely into primes: $\;2025$.

\item What is the remainder when $10^6$ is divided by $7$?

\item Find the last digit of $3^{2024}$.

\item For what single digit $d$ is the number $73d4$ divisible by $9$?

\item The positive integers $a$ and $b$ are distinct. Given $\gcd(a,b)=6$ and $\text{lcm}(a,b)=36$, find $a+b$.

\item Find the largest integer $k$ such that $2^k$ divides $20!$.

\item Find the remainder when $11^{10}$ is divided by $10$.

\item Find the remainder when $2^{100} + 3^{100}$ is divided by $5$.

\item What is the smallest prime factor of $1001$?

\item Find the only positive integer $n$ such that $n^2 - 1$ is a prime number.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Divisor counts, trailing zeros, modular congruence. Resist the urge to brute force.}

\begin{enumerate}

\item Find the number of positive divisors of $360$.

\item A positive integer $x$ leaves a remainder of $3$ when divided by $5$, and a remainder of $5$ when divided by $7$. Find the smallest possible value of $x$.

\item What is the sum of the digits of the number $2^{2024} \times 5^{2025}$?

\item Solve for the smallest positive integer $x$ such that $4x \equiv 5 \pmod 7$.

\item Find the last two digits of $99^{2}$.

\item Find the greatest common divisor of $2^{12}-1$ and $2^{18}-1$.

\item The sum of three consecutive positive integers is a perfect cube. Find the smallest possible value for the largest of these three integers.

\item Find the smallest positive integer $n$ such that $10n$ is a perfect square and $6n$ is a perfect cube.

\item Find the number of trailing zeros in $100!$.

\item The 5-digit number $24x8y$ is divisible by $4$ and by $9$. Find the maximum possible value of $x + y$.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Look for divisibility logic, algebraic structure, and bounds before computing.}

\begin{enumerate}

\item Find the sum of all integers $x$ such that $\dfrac{x^2 + 5}{x+1}$ is an integer.

\item Find the number of ordered pairs of positive integers $(a,b)$ such that $\dfrac{1}{a} + \dfrac{1}{b} = \dfrac{1}{6}$.

\item For how many integers $n \in [1, 100]$ is the polynomial $x^n - 1$ divisible by $x^2 + x + 1$?

\item The 5-digit number $A34B2$, where $A$ and $B$ are single digits and $A \neq 0$, is a multiple of $99$. Find the value of the product $A \times B$. \textit{\small(after SMC 2018 Q16)}

\item The product of 24 integers is equal to 1. Their sum is $0$. How many of the integers are equal to $-1$? \textit{\small(after SMC 2015 Q11)}

\item Find the largest integer $x$ such that $2^x$ divides $3^{256} - 1$.

\item Find all pairs of primes $(p,q)$ such that $p^2 - 2q^2 = 1$.

\item A palindromic 6-digit number $N$ is formed by placing the digits of a 3-digit number $x$ followed by the same digits in reverse order (for example, if $x=123$, $N=123321$). What is the largest prime number that must be a factor of all such numbers $N$? \textit{\small(after TMUA 2019 Paper 1 Q12)}

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(TMUA / SMC difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Insight over computation. Each question has a short, elegant solution --- find it.}

\begin{enumerate}

\item Find the sum of all integers $n$ such that $\dfrac{n^3+100}{n+10}$ is an integer.

\item How many 5-digit numbers can be formed using only the digits $1$, $2$, and $3$, such that the number is a multiple of $3$?

\item Let $P(x)$ be a polynomial with integer coefficients such that $P(1) = 10$ and $P(5) = 2$. Find all possible integer roots of $P(x)$.

\item Find the number of positive divisors of the smallest positive integer $n$ such that $\dfrac{n}{2}$ is a perfect square, $\dfrac{n}{3}$ is a perfect cube, and $\dfrac{n}{5}$ is a perfect fifth power.

\item Find all pairs of positive integers $(x,y)$ such that $x^2 - y! = 2016$.

\end{enumerate}

\SpeedClosing{``Speed comes from recognition, not from rushing.
Every shortcut is a pattern you have internalised.''}

\end{document}
