\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Number Theory}{7}

\begin{document}

\SpeedTitleBlock{Daily Number Theory Drill \#7}{AI Generation}
\noindent\textit{New toolkit today: Combinatorial Number Theory $\cdot$ Divisor Counting $\cdot$ Synoptic Mastery.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Primes, moduli, and basics. Pure recognition.}

\begin{enumerate}

\item Find the prime factorisation of $91$.

\item What is the last digit of $3^{20}$?

\item Find $\gcd(144, 60)$.

\item Compute $12 \times 13 \pmod{5}$.

\item What is the smallest prime factor of $323$?

\item How many positive divisors does $28$ have?

\item Compute $5^{3} \pmod{6}$.

\item Find $x \pmod{7}$ if $4x \equiv 1 \pmod{7}$.

\item What is the remainder when $10^{10}$ is divided by $9$?

\item Which is larger: $2^{30}$ or $3^{20}$?

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Index laws, modular equations, and divisor counting.}

\begin{enumerate}

\item Find the last two digits of $7^4$.

\item How many trailing zeros are there in $20!$?

\item Find the smallest positive integer $n$ such that $120n$ is a perfect square.

\item Find the largest power of $3$ that divides $50!$.

\item Find the unique pair of positive integers $(x,y)$ such that $x^2 - y^2 = 17$.

\item What is the remainder when $3^{100}$ is divided by $10$?

\item Solve for $x$ in the range $0 \leq x < 11$: $\;3x + 4 \equiv 6 \pmod{11}$.

\item Find the sum of the positive divisors of $50$.

\item How many prime numbers are there strictly between $40$ and $50$?

\item Find the smallest three-digit positive integer that leaves a remainder of $2$ when divided by $3, 4$, and $5$.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Synoptic links connecting number theory to algebra and combinatorics.}

\begin{enumerate}

\item Find the product of all positive integers $n < 50$ that have exactly three positive divisors. \textit{\small(after SMC 2015 Q16)}

\item Find the smallest positive integer $n$ such that $n$ has exactly $15$ positive divisors. \textit{\small(after SMC 2018 Q20)}

\item How many positive divisors of $3^5 \times 5^4 \times 7^3$ are multiples of $45$? \textit{\small(after SMC 2021 Q19)}

\item How many positive divisors of $10!$ are perfect squares?

\item What is the greatest possible value of $\gcd(n^2 + 5, n + 2)$ for a positive integer $n$?

\item Find the sum of all digits of the number $10^{15} - 2026$.

\item Find all pairs of positive integers $(x,y)$ such that $x^3 - y^3 = 37$. \textit{\small(after SMC 2023 Q1)}

\item Find the smallest positive integer $x$ such that $x \equiv 1 \pmod 3$, $x \equiv 2 \pmod 4$, and $x \equiv 3 \pmod 5$.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(TMUA / SMC difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Insight over computation. Each question has a short, elegant solution --- find it.}

\begin{enumerate}

\item How many integer pairs $(x,y)$ satisfy the equation $x^2 + xy + y^2 = 7$? \textit{\small(after SMC 2023 Q10)}

\item For how many integers $n \in \{1, 2, \dots, 100\}$ is $n^n$ a perfect square?

\item Find all pairs of positive integers $(a,b)$ such that $a^2 - 4b^2 = 45$. \textit{\small(after MAT 2021 Q1)}

\item What is the largest prime factor of $2^{16} - 1$?

\item If $p$ and $q$ are prime numbers such that $p^2 - 2q^2 = 1$, find the value of $p+q$. \textit{\small(after SMC 2023 Q17)}

\end{enumerate}

\SpeedClosing{``Speed comes from recognition, not from rushing.
Every shortcut is a pattern you have internalised.''}

\end{document}
