\documentclass[11pt,a4paper]{article}
\input{../../shared/preamble}
\SpeedHeader{Number Theory}{2}

\begin{document}

\SpeedTitleBlock{Daily Number Theory Drill \#2}{David Akinyele-Aje}
\noindent\textit{New toolkit today: Modular Arithmetic $\cdot$ Remainder Equations $\cdot$ Parity in Modulo.}

\vspace{2pt}
\noindent\textit{Attempt each question before checking answers. Time yourself. No calculators.}

\section*{Section A \quad Rapid Recognition \hfill{\normalfont\small($\leq$15\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Prime factors, parity, modular remainders. Pure recognition.}

\begin{enumerate}

\item Prime factorise $1001$.

\item What is the remainder when $10^5$ is divided by $11$?

\item Find the last digit of $7^{100}$.

\item Find the smallest positive integer $k$ such that $12k$ is a perfect square.

\item Find the greatest common divisor of $84$ and $120$.

\item What is the smallest prime $p > 30$?

\item Find the remainder when $3^{10}$ is divided by $8$.

\item Find the sum of all prime factors of $210$.

\item For what single digit $d$ is the number $53d2$ a multiple of $9$?

\item Find the largest integer $k$ such that $3^k$ divides $27^5$.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section B \quad Manipulation Drills \hfill{\normalfont\small($\leq$50\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Divisor counts, trailing zeros, modular congruence. Resist the urge to brute force.}

\begin{enumerate}

\item Find the number of positive divisors of $144$.

\item Find the smallest positive integer $x > 1$ that leaves a remainder of $1$ when divided by $3, 4,$ and $5$.

\item Solve for the smallest positive integer $x$ such that $5x \equiv 1 \pmod 7$.

\item Find the highest power of $5$ dividing $100!$.

\item Find the last two digits of $51^2$.

\item Find the smallest positive integer $n$ such that $n!$ is a multiple of $100$.

\item Given that $a$ and $b$ are distinct positive integers with $\gcd(a, b) = 12$ and $\text{lcm}(a, b) = 144$, find the minimum possible value of $a+b$.

\item What is the sum of the digits of the number $2^{10} \times 5^{8}$?

\item What is the remainder when $2^{20} + 3^{20}$ is divided by $5$?

\item Find the number of trailing zeros in $125!$.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section C \quad Substitution \& Structure \hfill{\normalfont\small($\leq$90\,s each)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Look for divisibility logic, algebraic structure, and bounds before computing.}

\begin{enumerate}

\item Find all positive integers $n$ such that $n^2 - 4n - 5$ is a prime number. \textit{\small(after SMC 2021 Q14)}

\item How many integers $x$ are there such that $\dfrac{x^2+11}{x-1}$ is an integer? \textit{\small(after SMC 2019 Q20)}

\item Find the sum of all positive integers $n$ such that $n^2 + 8n$ is a perfect square. \textit{\small(after BMO1 2018 Q1)}

\item How many 3-digit numbers have the property that the product of their digits is a prime number? \textit{\small(after SMC 2023 Q17)}

\item How many integers $x$ with $1 \le x \le 100$ satisfy $x^2 \equiv 4 \pmod 5$? \textit{\small(after SMC 2018 Q16)}

\item Find the number of ordered pairs of positive integers $(x,y)$ such that $x+y=100$ and $\gcd(x,y)=5$. \textit{\small(after SMC 2022 Q12)}

\item Find the largest integer $n$ such that $n^3 + 100$ is divisible by $n+10$. \textit{\small(after MAT 2018 Q1)}

\item Find the sum of all positive integers $x < 20$ such that $x^2 \equiv 5 \pmod{11}$. \textit{\small(after SMC 2019 Q22)}

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section*{Section D \quad Challenge \hfill{\normalfont\small(TMUA / SMC difficulty)}}
% ══════════════════════════════════════════════════════════════════════════════

\noindent\textit{Insight over computation. Each question has a short, elegant solution --- find it.}

\begin{enumerate}

\item A 4-digit number $A$ is formed by arranging the digits $1, 2, 3, 4$ in some order. What is the sum of all such possible numbers $A$? \textit{\small(after SMC 2021 Q21)}

\item Find the number of pairs of positive integers $(a,b)$ such that $a^2 - b^2 = 2024$. \textit{\small(after BMO1 2021 Q1)}

\item Let $S$ be the number of trailing zeros of $2025!$. What is the remainder when $S$ is divided by $100$? \textit{\small(after SMC 2023 Q12)}

\item How many pairs of positive integers $(x,y)$ with $1 \le x, y \le 100$ satisfy $x^2 + y^2 \equiv 0 \pmod 5$? \textit{\small(after BMO1 2017 Q2)}

\item Find the maximum value of $\gcd(n^2+3, n+2)$ as $n$ varies over all positive integers. \textit{\small(after BMO1 2019 Q2)}

\end{enumerate}

\SpeedClosing{``Speed comes from recognition, not from rushing.
Every shortcut is a pattern you have internalised.''}

\end{document}
